### 2016-12-20 Analysis/Character_difference/Modifying_matrices.Rmd
Cleaned Analysis folder

---
title: "Generating and modifying matrices"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 12
---

#### Loading the package and functions

```{r, eval = FALSE}
## The latest version of devtools
if(!require(devtools)) install.packages("devtools")

## The latest version of dispRity 
install_github("TGuillerme/dispRity")

## The latest version of Claddis
install_github("graemetlloyd/Claddis")

## The cran version of diversitree
if(!require(diversitree)) install.packages("diversitree")
```

```{r, results="hide", message=FALSE}
library(Claddis) ; library(dispRity) ; library(diversitree)
source("functions.R") ; load.functions(test = FALSE)

## Load the C libraries
dyn.load("../Functions/char.diff.so")

## Seed for repeatability
set.seed(0)
```

Generating the matrix
===============

Before anything, let's set our simulation variables: the number of taxa and the number of characters:

```{r}
## Number of taxa
ntaxa <- 25

## Number of characters
ncharacters <- 100
```

Generating the tree
-----------------------------------

First, we need to generate a realistic birth death tree upon which to simulate our characters.
For that we will need a function for randomly sampling our birth/death rates from a uniform (0,1) distribution while making sure that the death rate is lower that the birth rate:

```{r}
sample.birth.death <- function() {
    birth <- runif(min = 0, max = 1, 1)
    death <- runif(min = 0, max = birth, 1)
    return(c(birth, death))
}
```

Then let's generate the tree.
We add a while loop to the algorithm to make sure that the tree actually generate properly.
For example, if the birth rate sampled is 0, no tree will be generated!

```{r}
tree <- NULL
## The birth-death parameters
(birth_death <- sample.birth.death())
## Generate the tree
while(is.null(tree)) {
    tree <- tree.bd(birth_death, max.taxa = ntaxa)
}
```

We can then remove the tip labels (that conflict somehow with `phyclust` functions used downstream) and arbitrarily set the first species as the outgroup (thus rooting the tree).

```{r}
## Remove the $node labels (bugging for some reason)
tree$node.label <- NULL

## Make the first taxa (arbitrarily) the outgroup
tree$tip.label[[1]] <- "outgroup"
tree <- root(tree, "outgroup", resolve.root = TRUE)
```

And this is what our tree looks like

```{r, fig.width=6, fig.height=6}
plot(tree, cex = 0.8)

```
<!-- ```{r}
## Save the tree
simulationID = "0001"
write.nexus(tree, file = paste(ntaxa,"t_", ncharacters, "c_", simulationID, "_truetree.nex", sep = ""))
``` -->

Generating the matrix
-----------------------------------

We can now generate the matrix by using a `"mixed"` model (using both HKY-binary and Mk) with a state distribution of 0.85 binary characters and 0.15 three states characters and an evolutionary rate sampled from a gamma distribution with alpha = 5 and beta = 10.
We will also make sure there are no invariant characters and that the consistency index of the matrix is higher than 0.26.

```{r}
## Simulate a matrix
matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85,0.15),
    model = "mixed", rates = c(rgamma, rate = 100, shape = 3),
    invariant = FALSE)

## Resimulate it if the consistency index is too low
while(check.morpho(matrix_norm)[2,] < 0.26) {
    matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85,0.15),
        model = "mixed", rates = c(rgamma, rate = 100, shape = 3),
        invariant = FALSE)
}
```

We can also check how realistic our simulated matrix is:

```{r}
check.morpho(matrix_norm, tree)
```

Modifying the matrix
-----------------------------------

We can now maximise, minimise and randomise the character differences from the original matrix (`matrix_norm`).

```{r}
## Maximising the character difference
matrix_maxi <- modify.matrix(matrix_norm, type = "maximise", threshold = 0.25)
## Minimising the character difference
matrix_mini <- modify.matrix(matrix_norm, type = "minimise", threshold = 0.75)
## Randomising the character difference
matrix_rand <- modify.matrix(matrix_norm, type = "randomise")
```

We can then visualise the modifications by looking at the pairwise character differences:

```{r, fig.width=12, fig.height=6}
par(mfrow = c(1, 2))
## The character difference density (per character)
plot.char.diff.density(char.diff(matrix_norm), "Normal matrix")
## The character difference matrix (pairwise)
plot.cor.matrix(char.diff(matrix_norm), main = "")
```

```{r, fig.width=12, fig.height=6}
par(mfrow = c(1, 2))
## The character difference density (per character)
plot.char.diff.density(char.diff(matrix_maxi), "Maximised matrix")
## The character difference matrix (pairwise)
plot.cor.matrix(char.diff(matrix_maxi), main = "")
```

```{r, fig.width=12, fig.height=6}
par(mfrow = c(1, 2))
## The character difference density (per character)
plot.char.diff.density(char.diff(matrix_mini), "Minimised matrix")
## The character difference matrix (pairwise)
plot.cor.matrix(char.diff(matrix_mini), main = "")
```

```{r, fig.width=12, fig.height=6}
par(mfrow = c(1, 2))
## The character difference density (per character)
plot.char.diff.density(char.diff(matrix_rand), "Randomised matrix")
## The character difference matrix (pairwise)
plot.cor.matrix(char.diff(matrix_rand), main = "")
```