### 2017-04-25 Writing/supplementary/Sup2-SimParam.tex
Simulation parameters sup. mat

\documentclass[12pt,letterpaper]{article}
\usepackage{natbib}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{dcolumn}
\usepackage{mathtools}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\newcolumntype{d}[1]{D{.}{.}{#1}}

\pagenumbering{arabic}

%Pagination style and stuff
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}
%\bibpunct{(}{)}{;}{a}{}{,}

%---------------------------------------------
%
%       START
%
%---------------------------------------------

\begin{document}

%Running head
\begin{flushright}
Version dated: \today
\end{flushright}
\bigskip
\noindent RH: Characters correlation

\bigskip
\medskip
\begin{center}

\noindent{\Large \bf Effect of discrete character correlation on tree inference}
\bigskip
\noindent{\Large \bf Supplementary material 2 - Simulation parameters}

\noindent {\normalsize \sc Thomas Guillerme$^1$$^*$, and Martin D. Brazeau$^1$}\\
\noindent {\small \it 
$^1$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\}
\end{center}
\medskip
\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\

%Line numbering
% \modulolinenumbers[1]
% \linenumbers

\section{Morphological matrices simulations}

\subsection{``True'' Tree simulation - using diversitree \texttt{R} package v 0.9-8 \cite{fitzjohndiversitree2012}}
\begin{itemize}
  \item Model
  \begin{itemize}
    \item Birth-death tree (\texttt{diversitree::tree.bd})
  \end{itemize}
  \item Parameters
  \begin{itemize}
    \item Birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform distribution (0,1) with $\lambda > \mu$.
    \item Stopped when the tree reached 25, 75 or 150 taxa (extinct taxa excluded).
  \end{itemize}
\end{itemize}

\subsection{``Normal'' matrix simulation - using dispRity \texttt{R} package v 0.4 \cite{thomas_guillerme_2016_55646}}
\begin{itemize}
  \item Model
  \begin{itemize}
    \item Mk or HKY-binary selected randomly (\texttt{dispRity::sim.morpho})
  \end{itemize}
  \item Parameters
  \begin{itemize}
    \item States frequency: 0.85 binary characters and 0.15 three states (for Mk only)
    \item Transition/transversion ratio (for HKY-binary): 2 
    \item Character rate (for both models): gamma distribution with a rate $\alpha$ = 100 and a shape $\beta$ = 5.
    \item No invariant characters.
  \end{itemize}
\end{itemize}

\subsection{``Maximised'', ``Minimised'' and ``Null'' matrices}
\begin{itemize}
  \item Maximised
  \begin{itemize}
    \item Replacing the $n$ characters with a character difference (CD) to all other characters $< 0.25$.
  \end{itemize}

  \item Minimised
  \begin{itemize}
    \item Replacing the $m$ character with a CD to all other characters $> 0.75$.
  \end{itemize}

  \item Null
  \begin{itemize}
    \item Randomly replacing $\frac{n+m}{2}$ characters.
  \end{itemize}

\end{itemize}

\section{Tree inference parameters}

\subsection{Maximum parsimony: PAUP* version 4.0a151 \cite{swofford2001paup}}

\begin{itemize}
  \item Search settings
  \begin{itemize}
    \item Max. trees = 500 with auto-increase of 500.
    \item Heuristic search with random sequence addition replicated 10 times
  \end{itemize}
  \item Summary settings
  \begin{itemize}
    \item Summarising up to $1\times10^9$ trees in a strict majority rule consensus tree (cut off = 50).
  \end{itemize}
\end{itemize}

\subsection{Bayesian: MrBayes version 3.2.7 \cite{Ronquist2012mrbayes}}

\begin{itemize}
  \item Models
  \begin{itemize}
    \item M\textit{kv} + $\Gamma_4$
  \end{itemize}

  \item Priors
  \begin{itemize}
    \item variable rate prior with an exponential shape prior set to 0.5
  \end{itemize}

  \item MCMC
  \begin{itemize}
    \item Two runs
    \item Six chains per run
    \item Generations $<$ $1\times10^9$
    \item Sample frequency = 200
    \item ASDS diagnosis frequency = 10000
    \item ASDS $<$ 0.01
    \item ESS $>>$ 200
  \end{itemize}

  \item Summary settings
  \begin{itemize}
    \item Strict majority consensus tree (cut off = 50) with a relative burning of 25\%.
  \end{itemize}

\end{itemize}

\bibliographystyle{../sysbio}
\bibliography{../References}

\end{document}