### 2017-06-05 Analysis/EffectCorrelation.Rmd
Pipelined analysis

---
title: "Effect of character correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Effect of character correlation
===============

Before starting
===============

Loading the functions
-----------------------------------

Getting the functions

```{r, eval = TRUE}
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
```

Loading the data
-----------------------------------

Loading the tree comparisons from TreeCmp

```{r, message = FALSE}
## 25t_100c
bay_25t_100c <- read.TreeCmp("25t_100c", type = "Bayesian")
par_25t_100c <- read.TreeCmp("25t_100c", type = "Parsimony")

## 25t_350c
bay_25t_350c <- read.TreeCmp("25t_350c", type = "Bayesian")
par_25t_350c <- read.TreeCmp("25t_350c", type = "Parsimony")

## 25t_1000c
bay_25t_1000c <- read.TreeCmp("25t_1000c", type = "Bayesian")
par_25t_1000c <- read.TreeCmp("25t_1000c", type = "Parsimony")

## 75t_100c
bay_75t_100c <- read.TreeCmp("75t_100c", type = "Bayesian")
par_75t_100c <- read.TreeCmp("75t_100c", type = "Parsimony")

## 75t_350c
bay_75t_350c <- read.TreeCmp("75t_350c", type = "Bayesian")
par_75t_350c <- read.TreeCmp("75t_350c", type = "Parsimony")

## 75t_1000c 
bay_75t_1000c <- read.TreeCmp("75t_1000c", type = "Bayesian")
par_75t_1000c <- read.TreeCmp("75t_1000c", type = "Parsimony")

## 150t_100c
bay_150t_100c <- read.TreeCmp("150t_100c", type = "Bayesian")
par_150t_100c <- read.TreeCmp("150t_100c", type = "Parsimony")

## 150t_350c
bay_150t_350c <- read.TreeCmp("150t_350c", type = "Bayesian")
par_150t_350c <- read.TreeCmp("150t_350c", type = "Parsimony")

## 150t_1000c 
bay_150t_1000c <- read.TreeCmp("150t_1000c", type = "Bayesian")
par_150t_1000c <- read.TreeCmp("150t_1000c", type = "Parsimony")

```

Analysis the differences
===============

Normalise the score
-----------------------------------

Normalising the scores using NTS

```{r, message = FALSE}
## 25t_100c
bay_25t_100c <- NTS(bay_25t_100c, 25)
par_25t_100c <- NTS(par_25t_100c, 25)

## 25t_350c
bay_25t_350c <- NTS(bay_25t_350c, 25)
par_25t_350c <- NTS(par_25t_350c, 25)

## 25t_1000c
bay_25t_1000c <- NTS(bay_25t_1000c, 25)
par_25t_1000c <- NTS(par_25t_1000c, 25)

## Combining into a list
t25_list <- list(
    "c100" = list("bayesian" = bay_25t_100c,
                  "parsimony" = par_25t_100c
                  ),
    "c350" = list("bayesian" = bay_25t_350c,
                  "parsimony" = par_25t_350c
                  ),
    "c1000" = list("bayesian" = bay_25t_1000c,
                   "parsimony" = par_25t_1000c
                   )
                )

## 75t_100c
bay_75t_100c <- NTS(bay_75t_100c, 25)
par_75t_100c <- NTS(par_75t_100c, 25)

## 75t_350c
bay_75t_350c <- NTS(bay_75t_350c, 25)
par_75t_350c <- NTS(par_75t_350c, 25)

## 75t_1000c
bay_75t_1000c <- NTS(bay_75t_1000c, 25)
par_75t_1000c <- NTS(par_75t_1000c, 25)

## Combining into a list
t75_list <- list(
    "c100" = list("bayesian" = bay_75t_100c,
                  "parsimony" = par_75t_100c
                  ),
    "c350" = list("bayesian" = bay_75t_350c,
                  "parsimony" = par_75t_350c
                  ),
    "c1000" = list("bayesian" = bay_75t_1000c,
                   "parsimony" = par_75t_1000c
                   )
                )

## 150t_100c
bay_150t_100c <- NTS(bay_150t_100c, 25)
par_150t_100c <- NTS(par_150t_100c, 25)

## 150t_350c
bay_150t_350c <- NTS(bay_150t_350c, 25)
par_150t_350c <- NTS(par_150t_350c, 25)

## 150t_1000c
bay_150t_1000c <- NTS(bay_150t_1000c, 25)
par_150t_1000c <- NTS(par_150t_1000c, 25)

## Combining into a list
t150_list <- list(
    "c100" = list("bayesian" = bay_150t_100c,
                  "parsimony" = par_150t_100c
                  ),
    "c350" = list("bayesian" = bay_150t_350c,
                  "parsimony" = par_150t_350c
                  ),
    "c1000" = list("bayesian" = bay_150t_1000c,
                   "parsimony" = par_150t_1000c
                   )
                )

```

Plotting the results for 25 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the best tree (RF)
plot.results.single(t25_list, metric = 2, what = "best",
    main = "comparison to 'best' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the best tree (Triplets)
plot.results.single(t25_list, metric = 4, what = "best",
    main = "comparison to 'best' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the null tree (RF)
plot.results.single(t25_list, metric = 2, what = "null",
    main = "comparison to 'null' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the null tree (Triplets)
plot.results.single(t25_list, metric = 4, what = "null",
    main = "comparison to 'null' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

Plotting the results for 75 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the best tree (RF)
plot.results.single(t75_list, metric = 2, what = "best",
    main = "comparison to 'best' tree (75 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the best tree (Triplets)
plot.results.single(t75_list, metric = 4, what = "best",
    main = "comparison to 'best' tree (75 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the null tree (RF)
plot.results.single(t75_list, metric = 2, what = "null",
    main = "comparison to 'null' tree (75 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the null tree (Triplets)
plot.results.single(t75_list, metric = 4, what = "null",
    main = "comparison to 'null' tree (75 taxa)", legend = TRUE, NTS = TRUE)
```

Plotting the results for 150 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the best tree (RF)
plot.results.single(t150_list, metric = 2, what = "best",
    main = "comparison to 'best' tree (150 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the best tree (Triplets)
plot.results.single(t150_list, metric = 4, what = "best",
    main = "comparison to 'best' tree (150 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the null tree (RF)
plot.results.single(t150_list, metric = 2, what = "null",
    main = "comparison to 'null' tree (150 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the null tree (Triplets)
plot.results.single(t150_list, metric = 4, what = "null",
    main = "comparison to 'null' tree (150 taxa)", legend = TRUE, NTS = TRUE)
```

Checking the results for 25 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (RF)
plot.results.single(t25_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (Triplets)
plot.results.single(t25_list, metric = 4, what = "true",
    main = "comparison to 'true' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Checking comparisons (RF)
plot.results.single(t25_list, metric = 2, what = "check",
    main = "Check comparison (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Checking comparisons (Triplets)
plot.results.single(t25_list, metric = 4, what = "check",
    main = "Check comparison (25 taxa)", legend = TRUE, NTS = TRUE)
```

Checking the results for 75 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (RF)
plot.results.single(t75_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (75 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (Triplets)
plot.results.single(t75_list, metric = 4, what = "true",
    main = "comparison to 'true' tree (75 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Checking comparisons (RF)
plot.results.single(t75_list, metric = 2, what = "check",
    main = "Check comparison (75 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Checking comparisons (Triplets)
plot.results.single(t75_list, metric = 4, what = "check",
    main = "Check comparison (75 taxa)", legend = TRUE, NTS = TRUE)
```

Checking the results for 150 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (RF)
plot.results.single(t150_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (Triplets)
plot.results.single(t150_list, metric = 4, what = "true",
    main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Checking comparisons (RF)
plot.results.single(t150_list, metric = 2, what = "check",
    main = "Check comparison (150 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Checking comparisons (Triplets)
plot.results.single(t150_list, metric = 4, what = "check",
    main = "Check comparison (150 taxa)", legend = TRUE, NTS = TRUE)
```

Comparison to 'best' tree (RF normalised distance)

```{r, fig.width=12, fig.height=12}
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 2, what = "best",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 2, what = "best",
    ylab = "RF for 150 taxa", NTS = TRUE)
par(op)
```

Comparison to 'best' tree (Triplets normalised distance)

```{r, fig.width=12, fig.height=12}
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 4, what = "best",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 4, what = "best",
    ylab = "Tr for 150 taxa", NTS = TRUE)
par(op)
```

Comparison to 'null' tree (RF normalised distance)

```{r, fig.width=12, fig.height=12}
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 2, what = "null",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 2, what = "null", ylab = "RF for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 2, what = "null",
    ylab = "RF for 150 taxa", NTS = TRUE)
par(op)
```

Comparison to 'null' tree (Triplets normalised distance)

```{r, fig.width=12, fig.height=12}
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 4, what = "null",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 4, what = "null",
    ylab = "Tr for 150 taxa", NTS = TRUE)
par(op)
```