### 2017-07-17 Analysis/Character_difference/Comparing_characters.Rmd
Added formula in LaTeX

---
title: "Comparing characters"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 12
---

Before starting
===============

Loading the package
-----------------------------------

```{r, eval = TRUE}
library(dispRity) 
```

Comparing characters?
===============

The first step is to look at how to compare morphological characters.
How can we efficiently measure the correlation between two characters?
Of course, in some examples, it is pretty easy: a character `A {0,1,1,1,1}` is completely correlated to a character `B {0,1,1,1,1}`.
Moreover, `A {0,1,1,1,1}` must also be completely correlated with `C {1,0,0,0,0}`.
But for two other characters `D {0,0,1,1,1}` and `E {0,1,0,1,1}`, which one is more correlated to `A` (if any)?

Here we are going to explore the inverse absolute Spearman correlation (ranging from 0 - full correlation - to 1 - no correlation) as follows:

```{r}
cor.spearman <- function(X, Y) {
    return(1-round(abs(cor(X,Y, method = "spearman")), digit = 10))
}
```
> Note that to avoid rounding issues, we will round the correlation value up to the last 10 digits

We will also introduce a new difference called "Character Difference" that counts the difference in states pattern (ranging from 0 - no difference - to 1 - complete difference) as follows:

$$1 - \left( \frac{\mid {\frac{\abs{\sum_{i}^{n}(\mid x_{i} - y_{i}\mid )}} {n}-\frac{1}{2}} \mid  }{\frac{1}{2}}\right)$$

Where *n* is the number of taxa with comparable data and *x_i* and *y_i* are each characters states for the characters *x* and *y* and the taxon *i*.

Note that in practice, the characters are first normalised and the Gower distance is then calculated:
 * The normalisation works by changing the characters tokens labels to always start from `0` being the first state.
For example, a character `{8,8,2,1,3,0}` will be normalised by changing all the occurrences of the first state (`8`) by `0`, then all the occurrences of the second one (`2`) by `1` and so on resulting in the character being normalised as `{0,0,1,2,3,4}`.
Note that from a phylogenetic point of view, if the character is Fitch-like, both characters convey the exact same information.
 * The Gower distance is simply the distance divided by the number of observation (i.e. to take into account the inapplicable or missing data).

The function to calculate the character difference is a modification of the `stats::dist` function's source code (in `C`):

A simple case
----------------
Let's create a bunch of binary characters first:

```{r}
A <- c(1,0,0,0,0)
B <- c(0,1,1,1,1)
C <- c(1,1,0,0,0)
D <- c(0,1,1,0,0)
E <- c(1,0,0,1,1)
```

Straight away, we expect `A` and `B` to be totally correlated as well as `D` and `E`.
Also, we will check if our difference metric is triangular (i.e. if `A` = `B` then the difference `(A,X)` must be equal to `(B,X)`).

```{r}
## As expected:
cor.spearman(A,B)
cor.spearman(D,E)
## Note that the order doesn't matters
cor.spearman(B,A)
cor.spearman(E,D)
## And concerning the triangularity
cor.spearman(A,C) == cor.spearman(B,C)
cor.spearman(D,C) == cor.spearman(E,C)
```

This works as well for our character difference metric

```{r}
char.diff(matrix(c(A,B), nrow = 5))[2,1]
char.diff(matrix(c(D,E), nrow = 5))[2,1]
char.diff(matrix(c(B,A), nrow = 5))[2,1]
char.diff(matrix(c(E,D), nrow = 5))[2,1]
char.diff(matrix(c(A,C), nrow = 5))[2,1] ==
char.diff(matrix(c(B,C), nrow = 5))[2,1]
char.diff(matrix(c(D,C), nrow = 5))[2,1] ==
char.diff(matrix(c(E,C), nrow = 5))[2,1]
```

Let's have a look at the differences between `A` and the other characters:

```{r, fig.width=6, fig.height=6}
## Calculating the differences/correlations
character_correlations <- unlist(lapply(list(B,C,D,E), cor.spearman, A))
character_differences <- char.diff(matrix(c(A,B,C,D,E), nrow = 5))[1,-1]

## Plotting the results
plot(character_differences, type = "l", col = "blue", ylim = c(0,1), xaxt = "n", xlab = "",
    ylab = "", main = "Characters comparisons (binary)")
axis(1, at = 1:4, labels = c("A:B", "A:C", "A:D", "A:E"))
lines(character_correlations, col = "orange")
legend("topleft", legend = c("Difference", "Correlation"), lty = 1, xjust = 1, yjust = 1,
    col = c("blue", "orange"))
```

Although both metrics show different results, the magnitudes of changes seems consistent between them.
In other words, the "Character difference" seems to pick up slightly higher differences than the correlation one but overall the same trend is observed (increase differences and then no differences).

A non-binary case
----------------
However, in reality, some characters might have more than two state:

```{r}
A <- c(1,2,0,0,0)
B <- c(2,3,4,4,4)
C <- c(1,2,1,0,0)
D <- c(2,0,2,1,1)
E <- c(1,2,2,0,0)
F <- c(0,1,2,0,0)
G <- c(0,2,1,0,0)
```

Similarly, we expect `A` and `B` to be totally totally correlated as well as `C` and `D`.
However, the Spearman correlation does pick up some difference (i.e. not totally correlated)

```{r}
cor.spearman(A,B)
cor.spearman(C,D)
## And triangularity is not conserved!
cor.spearman(C,F) == cor.spearman(D,F)
```
Thankfully though, the "Character Difference" metric has no problems!

```{r}
char.diff(matrix(c(A,B), nrow = 5))[2,1]
char.diff(matrix(c(C,D), nrow = 5))[2,1]
char.diff(matrix(c(A,F), nrow = 5))[2,1] ==
char.diff(matrix(c(B,F), nrow = 5))[2,1]
char.diff(matrix(c(C,F), nrow = 5))[2,1] ==
char.diff(matrix(c(D,F), nrow = 5))[2,1]
```

Let's have a look at the differences between `A` and the other characters:

```{r, fig.width=6, fig.height=6}
## Calculating the differences/correlations
character_correlations <- unlist(lapply(list(B,C,D,E,F,G), cor.spearman, A))
character_differences <- char.diff(matrix(c(A,B,C,D,E,F,G), nrow = 5))[1,-1]

## Plotting the results
plot(character_differences, type = "l", col = "blue", ylim = c(0,1), xaxt = "n", xlab = "",
    ylab = "", main = "Characters comparisons (non-binary)")
axis(1, at = 1:6, labels = c("A:B", "A:C", "A:D", "A:E", "A:F", "A:G"))
lines(character_correlations, col = "orange")
legend("topleft", legend = c("Difference", "Correlation"), lty = 1, xjust = 1, yjust = 1,
    col = c("blue", "orange"))
```

The correlation metric does not match the actual differences between characters anymore!

Some more complex example
===============

Making up some data
-----------------------------------
First let's make up some discrete morphological matrices using the `dispRity` package.

```{r}
set.seed(0)
# A random tree (coalescent)
tree <- rcoal(15)
# A morphological matrix (Mk model)
matrix <- sim.morpho(tree, 50, states = c(0.85, 0.15), rates = c(rgamma, rate = 10, shape = 5),
    invariant = FALSE)
# Checking matrix
check.morpho(matrix, tree)

## This produces a fairly "realistic" matrix
# Transforming the characters to be numeric
matrix <- apply(matrix, 2, as.numeric)
```

Let's now run the pairwise comparisons for both metrics. We will use the `cor.matrix` function for calculating the spearman correlation (slow).

```{r}
source("../../Functions/cor.matrix.R")
pairwise_matrix_correlation <- cor.matrix(matrix, cor.spearman)
pairwise_matrix_differences <- char.diff(matrix)
```

We can now visualise a same single pairwise comparison (character `1` *versus* the others):

```{r, fig.width=6, fig.height=6}
source("../../Functions/plot.cor.matrix.R")
## Extracting the first character differences
character_differences <- pairwise_matrix_differences[,1]
character_correlation <- pairwise_matrix_correlation[,1]

## Sorting the values following the characters differences
sorting <- sort(character_differences, index.return = TRUE)[[2]]
character_differences <- character_differences[sorting]
character_correlation <- character_correlation[sorting]

## Plotting the results
plot(character_differences, type = "l", col = "blue", ylim = c(0,1), xlab = "", ylab = "",
    main = "Characters comparisons (non-binary)")
lines(character_correlation, col = "orange")
legend("topleft", legend = c("Difference", "Correlation"), lty = 1, xjust = 1, yjust = 1,
    col = c("blue", "orange"))
```
Note how the correlations quickly seem to plateau around 0.6-0.8 with a decrease in variance as the character difference increase (see below for replicated simulations).

We can also visualise the whole pairwise character differences/correlation distributions:

```{r, fig.width = 12, fig.height = 12}
op <- par(mfrow = c(2,2), bty = "n")
plot.cor.matrix(pairwise_matrix_differences, main = "Pairwise Characters Differences")
plot.cor.matrix(pairwise_matrix_correlation, main = "Pairwise Correlations score")
hist(na.omit(as.vector(pairwise_matrix_differences)), breaks = 15,
    main = "Differences distribution", xlab = "Character differences")
hist(na.omit(as.vector(pairwise_matrix_correlation)), breaks = 15,
    main = "Correlations distribution", xlab = "Character correlations")
par(op)
```

Two interesting points to note for the character differences metric:
 * The metric is not continuous: some values are never achieved!
 * The distribution looks like "half" of a normal distribution. This is easily explainable since we consider "mirror" character (e.g. `A {0,1,1}` and `B {1,0,0}`) to be identical and thus only one side of the distribution is explored. 
We can also have a look at the difference between both methods.
Ideally this difference (the residuals) should be normally distributed.

```{r, fig.width = 6, fig.height = 6}
## Calculating the differences
pairwise_differences <- pairwise_matrix_differences-pairwise_matrix_correlation
## Plotting the differences
hist(na.omit(as.vector(pairwise_differences)), breaks = 15,
    main = "Methods Differences distribution", xlab = "Residuals")

## Are the residuals normal?
shapiro.test(na.omit(pairwise_differences))
```

Appendix
===============

Some more simulations
-----------------------------------
Because here we are affected by the random seed and the size of the matrix, here is a longer simulation running past more matrices:

For speed improvement, we will "functionise" the different procedures (i.e. making matrices, etc...).

```{r}
# Function for looping the first step (matrix generation)
make.a.matrix <- function(taxa = 80, characters = 400) {
    ## Generate a tree
    tree <- rcoal(taxa)
    ## Generate a matrix
    matrix <- sim.morpho(tree, characters, states = c(0.85, 0.15),
        rates = c(rgamma, rate = 10, shape = 5), invariant = FALSE, verbose = FALSE)
    # Make the matrix output numeric
    matrix <- apply(matrix, 2, as.numeric)

    return(matrix)
}
```

```{r}
# Number of replicates:
replicates = 100
# Generate the matrices
matrices <- replicate(replicates, make.a.matrix(), simplify = FALSE)

# Get the character differences
character_differences <- lapply(matrices, char.diff)
character_correlation <- lapply(matrices, cor.matrix, cor.spearman)

# Calculating the residuals
characters_differences_distribution <- NULL
characters_correlation_distribution <- NULL
for (replicate in 1:replicates) {
    characters_differences_distribution <- c(characters_differences_distribution,
        as.vector(na.omit(character_differences[[replicate]])))
    characters_correlation_distribution <- c(characters_correlation_distribution,
        as.vector(na.omit(character_correlation[[replicate]])))
}
residuals <- characters_differences_distribution - characters_correlation_distribution

# Sorting the first characters
first_character_differences <- matrix(ncol = replicates, nrow = nrow(character_differences[[1]]))
first_character_correlation <- first_character_differences
for (replicate in 1:replicates) {
    sorting <- sort(character_differences[[replicate]][,1], index.return = TRUE)[[2]]
    first_character_differences[,replicate] <- character_differences[[replicate]][,1][sorting]
    first_character_correlation[,replicate] <- character_correlation[[replicate]][,1][sorting]
}

# Calculating the medians of the first characters differences
median_first_character_differences <- apply(first_character_differences, 1, median)
median_first_character_correlation <- apply(first_character_correlation, 1, median)
```

Now let's plot all these results:

```{r, fig.width = 12, fig.height = 12}
op <- par(mfrow = c(2,2), bty = "n")
hist(characters_differences_distribution, breaks = 40, main = "Characters Difference distribution",
    xlab = "Characters Differences")
hist(characters_correlation_distribution, breaks = 40, main = "Characters Correlation distribution",
    xlab = "Characters Correlation")
plot(density(residuals), main = "Difference between both methods",
    xlab = "Residuals")
par(op)
```

And finally the example for the first character again:

```{r, fig.width = 10, fig.height = 10}
plot(first_character_differences[,1], type = "l", col = "lightblue", ylim = c(0,1), xlab = "",
    ylab = "", main = "Characters comparison (matrix 80x400)")
lines(first_character_correlation[,1], col = "yellow", lty = 3)
for(replicate in 2:replicates) {
    lines(first_character_differences[,replicate], col = "lightblue")
    lines(first_character_correlation[,replicate], col = "yellow", lty = 3)
}
# Adding the medians
lines(median_first_character_differences, col = "blue")
lines(median_first_character_correlation, col = "orange")
legend("topleft", legend = c("Differences", "Correlation"), lty = 1, xjust = 1,
    yjust = 1, col = c("blue", "orange"))
```

Digression on missing data
-----------------------------------

In practice some (all?) matrices contain a number of missing data entries. In the character difference metric, because we use the Gower distance, only comparable sites are taken into account.
For example, the difference between two characters `A = {0,0,?,?,?}` and `B = {0,1,1,?,?}` will be obtained by comparing only `A' = {0,0}` and `B' = {0,1}` (resulting into a difference of 1).
If a third character `C = {?,?,0,0,1}`, the difference between `A` and `C` is typically inapplicable (`NA`).