### 2017-07-26 Analysis/EffectCorrelationStats.Rmd
Separated stats bit for supplementaries

---
title: "Effect of character correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Effect of character correlation
===============

Loading the functions and the data
-----------------------------------

Getting the functions and the data

```{r, eval = TRUE, message = FALSE}
library(dispRity)
library(lme4)
library(lmerTest)
library(ordinal)
library(MASS)
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")
```

Nested anova
-----------------------------------

Let's analysis the data using a nested anova.
The models look at the effect of our scenarios ("maximised", "minimised", etc...) on the normalised tree score (NTS - using both the Robinson-Fould and Triplets scores) with a nested effect of the method (parsimony or bayesian), the number of taxa (25, 75, 150) and the number of characters (100, 150, 1000).

```{r}
## Getting the whole data
whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)

## getting the aov data
aov_data_RF_best <- get.aov.data(whole_data, "RF", "norm", combined = TRUE)
aov_data_Tr_best <- get.aov.data(whole_data, "Triples", "norm", combined = TRUE)
aov_data_RF_null <- get.aov.data(whole_data, "RF", "null", combined = TRUE)
aov_data_Tr_null <- get.aov.data(whole_data, "Triples", "null", combined = TRUE)

## Models with normalised tree score function of scenario (with nestedness)
model_RF_best <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_RF_best)
model_Tr_best <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_Tr_best)
model_RF_null <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_RF_null)
model_Tr_null <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_Tr_null)

## Testing model assumptions
## Normality of the residuals
norm_RF_best <- shapiro.test(residuals(model_RF_best))
norm_Tr_best <- shapiro.test(residuals(model_Tr_best))
norm_RF_null <- shapiro.test(residuals(model_RF_null))
norm_Tr_null <- shapiro.test(residuals(model_Tr_null))
normality <- list(norm_RF_best, norm_Tr_best, norm_RF_null, norm_Tr_null)

## Variance homoscedasticity
homo_RF_best <- bartlett.test(NTS ~ scenario, data = aov_data_RF_best)
homo_Tr_best <- bartlett.test(NTS ~ scenario, data = aov_data_Tr_best)
homo_RF_null <- bartlett.test(NTS ~ scenario, data = aov_data_RF_null)
homo_Tr_null <- bartlett.test(NTS ~ scenario, data = aov_data_Tr_null)
homoscedasticity <- list(homo_RF_best, homo_Tr_best, homo_RF_null, homo_Tr_null)

## Summarise the results
results <- matrix(nrow = 4, data = c(
                round(unlist(lapply(normality, `[[`, 1)), digit = 3),
                round(unlist(lapply(normality, `[[`, 2)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 1)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 2)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 3)), digit = 3)),
            dimnames = list(
                c("RF best", "Triples best", "RF null", "Triples null"),
                c("W", "p.value", "Bartlett's K-squared", "df", "p.value")))

## Check assumption's validities
validity <- apply(results, 1,
    function(X) ifelse((X[2] >= 0.05 && X[5] >= 0.05), TRUE, FALSE))

## Report results
(results <- cbind(as.data.frame(results), validity))

```

All our models violate a least one of the two models assumptions (normality of the residuals and homoscedasticity of variance) and thus cannot be used to compare the scenarios.

Distributions probability overlap (Bhattacharrya Coefficient)
-----------------------------------

We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don't overlap (significant effect of one scenario on topology).
Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).

```{r}
## Running the pairwise Bhattacharrya Coefficient analysis:
(bc_RF_best <- run.bhatt.coeff(whole_data, "RF", "norm"))
(bc_Tr_best <- run.bhatt.coeff(whole_data, "Triples", "norm"))
(bc_RF_null <- run.bhatt.coeff(whole_data, "RF", "null"))
(bc_Tr_null <- run.bhatt.coeff(whole_data, "Triples", "null"))
```

Non-parametric anova
-----------------------------------

We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don't overlap (significant effect of one scenario on topology).
Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).

```{r}
# ## Cumulative Link Mixed Models
# mm1 <- clmm(NTS ~ scenario + (1|taxa/method) + (1|character/method),
#             data = aov_data_RF_best)
# mm1
# summary(mm1)

# ## Ordinated logistic regression
# polr(scenario ~ NTS + (1) ,
#                       data = aov_data_RF_best)

```