### 2017-09-19 Analysis/EffectCorrelationStats.Rmd
Updated stats

---
title: "Effect of character correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Effect of character correlation

## Loading the functions and the data

Getting the functions and the data.

```{r, eval = TRUE, message = FALSE}
library(dispRity)
library(lme4)
library(lmerTest)
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")

## Getting the whole data
whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)
```

```{r, echo = FALSE}
library(knitr)
library(xtable)
```

# Pooled results

## Effect of character correlation per scenario

```{r, warning = FALSE}
## Pooling the data per scenarios
scenario_RF_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
                               metric = "RF", best = "norm")
scenario_Tr_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
                               metric = "Tr", best = "norm")

## Preparing the table labels
metric.labels <- c("RF", "", "", "Tr", "", "")

## Combining the data into a table
scenarios <- pool.table(list(scenario_RF_norm, scenario_Tr_norm),  metric.label = metric.labels,
                        label.param = "scenario")
```

```{r, echo = FALSE}
## Setting the table captions
captions_pooled_scenarios <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled scenarios.")

## Printing the tables
kable(scenarios, row.names = FALSE, caption = captions_pooled_scenarios)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios")
```

### Testing the effect of the scenario on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ scenario, fun = lm)
## Removing the null tests
results[[1]] <- results[[1]][-c(3,4),]
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
scenario_list <- list(scenario_RF_norm, scenario_Tr_norm)
metric.labels <- c("RF", "", "", "Tr", "", "")
test_table <- pair.test.table(scenario_list, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_scen_test <- paste("Difference between the pooled scenarios. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_scen_test)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test")
```

## Effect of character correlation per number of character

```{r, warning = FALSE}
## Pooling the data per character
character_RF_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
                               metric = "RF", best = "norm")
character_Tr_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
                               metric = "Tr", best = "norm")

## Preparing the table labels
metric.labels <- c("RF", "", "", "Tr", "", "")

## Combining the data into a table
characters <- pool.table(list(character_RF_norm, character_Tr_norm), metric.label = metric.labels,
                         label.param = "character")
```

```{r, echo = FALSE}
## Setting the table captions
captions_pooled_characters <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of characters.")

## Printing the tables
kable(characters, row.names = FALSE, caption = captions_pooled_characters)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(characters, caption = captions_pooled_characters, label = "Tab_pooledcharacters")
```

### Testing the effect of the number of characters on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ character, fun = lm)
## Removing the null tests
results[[1]] <- results[[1]][-c(3,4),]
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
character_list <- list(character_RF_norm, character_Tr_norm)
metric.labels <- c("RF", "", "", "Tr", "", "")
test_table <- pair.test.table(character_list, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_char_test <- paste("Difference between the pooled number of characters. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_char_test)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_char_test, label = "Tab_pooledscharacters_test")
```

## Effect of character correlation per number of taxa

```{r, warning = FALSE}
## Pooling the data per taxa
taxa_RF_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
                               metric = "RF", best = "norm")
taxa_Tr_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
                               metric = "Tr", best = "norm")

## Preparing the table labels
metric.labels <- c("RF", "", "", "Tr", "", "")

## Combining the data into a table
taxa <- pool.table(list(taxa_RF_norm, taxa_Tr_norm), metric.label = metric.labels,
                   label.param = "taxa")
```

```{r, echo = FALSE}
## Setting the table captions
captions_pooled_taxa <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of taxa.")

## Printing the tables
kable(taxa, row.names = FALSE, caption = captions_pooled_taxa)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(taxa, caption = captions_pooled_taxa, label = "Tab_pooledtaxa")
```

### Testing the effect of the method on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ taxa, fun = lm)
## Removing the null tests
results[[1]] <- results[[1]][-c(3,4),]
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
taxa_list <- list(taxa_RF_norm, taxa_Tr_norm)
metric.labels <- c("RF", "", "", "Tr", "", "")
test_table <- pair.test.table(taxa_list, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_tax_test <- paste("Difference between the pooled number of taxa. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_tax_test)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_tax_test, label = "Tab_pooledstaxa_test")
```

## Effect of character correlation per methods

```{r, warning = FALSE}
## Pooling the data per taxa
method_RF_norm <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
                               metric = "RF", best = "norm")
method_Tr_norm <- multi.pool(whole_data, param =c("bayesian", "parsimony"),
                               metric = "Tr", best = "norm")

## Preparing the table labels
metric.labels <- c("RF", "", "Tr", "")

## Combining the data into a table
methods <- pool.table(list(method_RF_norm, method_Tr_norm), metric.label = metric.labels,
                      label.param = "method")
```

```{r, echo = FALSE}
## Setting the table captions
captions_pooled_methods <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled methods.")

## Printing the tables
kable(methods, row.names = FALSE, caption = captions_pooled_methods)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(methods, caption = captions_pooled_methods, label = "Tab_pooledmethod")
```

### Testing the effect of the method on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ method, fun = lm)
## Removing the null tests
results[[1]] <- results[[1]][-c(3,4),]
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
methods_list <- list(method_RF_norm, method_Tr_norm)
metric.labels <- c("RF", "", "Tr", "")
test_table <- pair.test.table(methods_list, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_met_test <- paste("Difference between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_met_test)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_met_test, label = "Tab_pooledsmethods_test")
```

# Combined results

We can also analyse the effect of each parameters combination separately:

## Summary statistics

We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
The p-values are adjusted for multiple tests using Bonferroni-Holm correction.

```{r}
## Summarising the data
sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL)
```

```{r, echo = FALSE}
## Setting the table captions
captions_summary <- paste("Summary statistics of the normalised", rep(c("Robinson-Fould", "Triplets"), 2), "distance to the", c(rep("best", 2), rep("null", 2)), "tree.")

## Printing the tables
kable(sum_RF_best, row.names = FALSE, caption = captions_summary[1])
kable(sum_Tr_best, row.names = FALSE, caption = captions_summary[2])
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

## Tables
make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1])
make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2])
```

## Nested anova

Let's analysis the data using a nested anova.
The models look at the effect of our scenarios ("maximised", "minimised", etc...) on the normalised tree score (NTS - using both the Robinson-Fould and Triplets scores) with a nested effect of the method (parsimony or bayesian), the number of taxa (25, 75, 150) and the number of characters (100, 150, 1000).

```{r}
## Applying an lm to all the data with error
results <- test.effect.aov(whole_data, fun = lmer,
                           formula = NTS ~ scenario + (1|method) + (1|taxa) + (1|character))
## Removing the null tests
results[[1]] <- results[[1]][-c(3,4),]
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

All our models violate a least one of the two models assumptions (normality of the residuals and homoscedasticity of variance) and thus cannot be used to compare the scenarios.
This is due to the fact that the metric is bounded to 1.

## Distributions probability overlap (Bhattacharrya Coefficient)

We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don t overlap (significant effect of one scenario on topology).
Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).

```{r}
## Running the pairwise Bhattacharrya Coefficient analysis:
bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff)
bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff)

## Tables
kable(bc_RF_best, row.names = FALSE, digit = 3)
kable(bc_Tr_best, row.names = FALSE, digit = 3)
```

```{r, echo = FALSE}
## Setting the table captions
captions_bc <- paste0("Bhattacharrya Coefficient (",rep(c("Robinson-Fould distance", "Triplets distance"), 1) ," from the ",c(rep("best", 2)), " tree)" , " between the different scenarios.")

## Printing the tables
kable(bc_RF_best, row.names = FALSE, caption = captions_bc[1], digit = 3)
kable(bc_Tr_best, row.names = FALSE, caption = captions_bc[2], digit = 3)
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_bc <- paste0("Tab_BC", rep(c("RF", "Tr")), c(rep("best", 2)))

## Tables
make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1])
make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2])
```

## Non-parametric pairwise comparisons

We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
The p-values are adjusted for multiple tests using Bonferroni-Holm correction.

```{r, warning = FALSE}
## Running the pairwise Wilcox tests:
wx_RF_best <- apply.test(whole_data, "RF", "norm", wilcox.test)
wx_Tr_best <- apply.test(whole_data, "Triples", "norm", wilcox.test)

## Applying Bonferroni correction
for(column in c(5,7,9)) {
  wx_RF_best[,column] <- p.adjust(wx_RF_best[,column], method = "bonferroni")
  wx_Tr_best[,column] <- p.adjust(wx_Tr_best[,column], method = "bonferroni")
}
```

```{r, echo = FALSE}
## Setting the table captions
captions_wx <- paste0("Non-parametric Wilcoxon test (",rep(c("Robinson-Fould distance", "Triplets distance")) ," from the ",c(rep("best", 2)), " tree)" , " between the different scenarios (p-values corrected using Bonferroni-Holm correction).")

## Printing the tables
kable(wx_RF_best, row.names = FALSE, caption = captions_wx[1], digit = 3)
kable(wx_Tr_best, row.names = FALSE, caption = captions_wx[2], digit = 3)
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_wx <- paste0("Tab_BC", rep(c("RF", "Tr")), c(rep("best", 2)))

## Tables
make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1])
make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2])
```

# Pooled results

## Effect of character correlation per scenario

```{r, warning = FALSE}
## Pooling the data per scenarios
scenario_RF_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
                               metric = "RF", best = "norm")
scenario_Tr_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
                               metric = "Tr", best = "norm")
scenario_RF_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
                               metric = "RF", best = "rand")
scenario_Tr_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
                               metric = "Tr", best = "rand")

## Preparing the table labels
comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)

## Combining the data into a table
scenarios <- pool.table(list(scenario_RF_norm, scenario_Tr_norm, scenario_RF_null, scenario_Tr_null),
                    comp.label = comp.labels, metric.label = metric.labels, label.param = "scenario")
```

```{r, echo = FALSE}
library(knitr)
## Setting the table captions
captions_pooled_scenarios <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled scenarios.")

## Printing the tables
kable(scenarios, row.names = FALSE, caption = captions_pooled_scenarios)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
library(xtable)
make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios")
```

### Testing the effect of the scenario on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ scenario, fun = lm)
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
scenario_list <- list(scenario_RF_norm, scenario_Tr_norm, scenario_RF_null, scenario_Tr_null)
comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
test_table <- pair.test.table(scenario_list, comp.label = comp.labels, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_scen_test <- paste("Difference between the pooled scenarios. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_scen_test)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test")
```

## Effect of character correlation per number of character

```{r, warning = FALSE}
## Pooling the data per character
character_RF_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
                               metric = "RF", best = "norm")
character_Tr_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
                               metric = "Tr", best = "norm")
character_RF_null <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
                               metric = "RF", best = "rand")
character_Tr_null <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
                               metric = "Tr", best = "rand")

## Preparing the table labels
comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)

## Combining the data into a table
characters <- pool.table(list(character_RF_norm, character_Tr_norm, character_RF_null, character_Tr_null),
                    comp.label = comp.labels, metric.label = metric.labels, label.param = "character")
```

```{r, echo = FALSE}
## Setting the table captions
captions_pooled_characters <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of characters.")

## Printing the tables
kable(characters, row.names = FALSE, caption = captions_pooled_characters)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(characters, caption = captions_pooled_characters, label = "Tab_pooledcharacters")
```

### Testing the effect of the number of characters on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ character, fun = lm)
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
character_list <- list(character_RF_norm, character_Tr_norm, character_RF_null, character_Tr_null)
comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
test_table <- pair.test.table(character_list, comp.label = comp.labels, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_char_test <- paste("Difference between the pooled number of characters. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_char_test)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_char_test, label = "Tab_pooledscharacters_test")
```

## Effect of character correlation per number of taxa

```{r, warning = FALSE}
## Pooling the data per taxa
taxa_RF_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
                               metric = "RF", best = "norm")
taxa_Tr_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
                               metric = "Tr", best = "norm")
taxa_RF_null <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
                               metric = "RF", best = "rand")
taxa_Tr_null <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
                               metric = "Tr", best = "rand")

## Preparing the table labels
comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)

## Combining the data into a table
taxa <- pool.table(list(taxa_RF_norm, taxa_Tr_norm, taxa_RF_null, taxa_Tr_null),
                    comp.label = comp.labels, metric.label = metric.labels, label.param = "taxa")
```

```{r, echo = FALSE}
## Setting the table captions
captions_pooled_taxa <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of taxa.")

## Printing the tables
kable(taxa, row.names = FALSE, caption = captions_pooled_taxa)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(taxas, caption = captions_pooled_taxas, label = "Tab_pooledtaxa")
```

### Testing the effect of the method on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ taxa, fun = lm)
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
taxa_list <- list(taxa_RF_norm, taxa_Tr_norm, taxa_RF_null, taxa_Tr_null)
comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
test_table <- pair.test.table(taxa_list, comp.label = comp.labels, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_tax_test <- paste("Difference between the pooled number of taxa. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_tax_test)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_tax_test, label = "Tab_pooledstaxa_test")
```

## Effect of character correlation per methods

```{r, warning = FALSE}
## Pooling the data per taxa
method_RF_norm <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
                               metric = "RF", best = "norm")
method_Tr_norm <- multi.pool(whole_data, param =c("bayesian", "parsimony"),
                               metric = "Tr", best = "norm")
method_RF_null <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
                               metric = "RF", best = "rand")
method_Tr_null <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
                               metric = "Tr", best = "rand")

## Preparing the table labels
comp.labels <- c("Best", rep("", 3), "Null", rep("", 3))
metric.labels <- rep(c("RF", "", "Tr", ""), 2)

## Combining the data into a table
methods <- pool.table(list(method_RF_norm, method_Tr_norm, method_RF_null, method_Tr_null),
                    comp.label = comp.labels, metric.label = metric.labels, label.param = "method")
```

```{r, echo = FALSE}
## Setting the table captions
captions_pooled_methods <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled methods.")

## Printing the tables
kable(methods, row.names = FALSE, caption = captions_pooled_methods)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(methods, caption = captions_pooled_methods, label = "Tab_pooledmethod")
```

### Testing the effect of the method on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ method, fun = lm)
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
methods_list <- list(method_RF_norm, method_Tr_norm, method_RF_null, method_Tr_null)
comp.labels <- c("Best", rep("", 3), "Null", rep("", 3))
metric.labels <- rep(c("RF", "", "Tr", ""), 2)
test_table <- pair.test.table(methods_list, comp.label = comp.labels, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_met_test <- paste("Difference between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_met_test)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_met_test, label = "Tab_pooledsmethods_test")
```

# Individual results

We can also analyse the effect of each parameters combination separately:

## Summary statistics

We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
The p-values are adjusted for multiple tests using Bonferroni-Holm correction.

```{r}
## Summarising the data
sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL)
sum_RF_null <- apply.test(whole_data, "RF", "null", NULL)
sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL)
```

```{r, echo = FALSE}
## Setting the table captions
captions_summary <- paste("Summary statistics of the normalised", rep(c("Robinson-Fould", "Triplets"), 2), "distance to the", c(rep("best", 2), rep("null", 2)), "tree.")

## Printing the tables
kable(sum_RF_best, row.names = FALSE, caption = captions_summary[1])
kable(sum_Tr_best, row.names = FALSE, caption = captions_summary[2])
kable(sum_RF_null, row.names = FALSE, caption = captions_summary[3])
kable(sum_Tr_null, row.names = FALSE, caption = captions_summary[4])
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

## Tables
make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1])
make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2])
make.xtable(sum_RF_null, caption = captions_summary[3], label = labels_summary[3])
make.xtable(sum_Tr_null, caption = captions_summary[4], label = labels_summary[4])
```

## Nested anova

Let's analysis the data using a nested anova.
The models look at the effect of our scenarios ("maximised", "minimised", etc...) on the normalised tree score (NTS - using both the Robinson-Fould and Triplets scores) with a nested effect of the method (parsimony or bayesian), the number of taxa (25, 75, 150) and the number of characters (100, 150, 1000).

```{r}
## Applying an lm to all the data with error
results <- test.effect.aov(whole_data,
  formula = NTS ~ scenario + (1|method) + (1|taxa) + (1|character), fun = lmer)
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

All our models violate a least one of the two models assumptions (normality of the residuals and homoscedasticity of variance) and thus cannot be used to compare the scenarios.
This is due to the fact that the metric is bounded to 1.

## Distributions probability overlap (Bhattacharrya Coefficient)

We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don t overlap (significant effect of one scenario on topology).
Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).

```{r}
## Running the pairwise Bhattacharrya Coefficient analysis:
bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff)
bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff)
bc_RF_null <- apply.test(whole_data, "RF", "null", bhatt.coeff)
bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff)

## Tables
kable(bc_RF_best, row.names = FALSE, digit = 3)
kable(bc_RF_best, row.names = FALSE, digit = 3)
kable(bc_RF_best, row.names = FALSE, digit = 3)
kable(bc_RF_best, row.names = FALSE, digit = 3)
```

```{r, echo = FALSE}
## Setting the table captions
captions_bc <- paste0("Bhattacharrya Coefficient (",rep(c("Robinson-Fould distance", "Triplets distance"), 2) ," from the ",c(rep("best", 2), rep("null", 2)), " tree)" , " between the different scenarios.")

## Printing the tables
kable(bc_RF_best, row.names = FALSE, caption = captions_bc[1], digit = 3)
kable(bc_Tr_best, row.names = FALSE, caption = captions_bc[2], digit = 3)
kable(bc_RF_null, row.names = FALSE, caption = captions_bc[3], digit = 3)
kable(bc_Tr_null, row.names = FALSE, caption = captions_bc[4], digit = 3)
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_bc <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

## Tables
make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1])
make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2])
make.xtable(bc_RF_null, caption = captions_bc[3], label = labels_bc[3])
make.xtable(bc_Tr_null, caption = captions_bc[4], label = labels_bc[4])
```

## Non-parametric pairwise comparisons

We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
The p-values are adjusted for multiple tests using Bonferroni-Holm correction.

```{r, warning = FALSE}
## Running the pairwise Wilcox tests:
wx_RF_best <- apply.test(whole_data, "RF", "norm", wilcox.test)
wx_Tr_best <- apply.test(whole_data, "Triples", "norm", wilcox.test)
wx_RF_null <- apply.test(whole_data, "RF", "null", wilcox.test)
wx_Tr_null <- apply.test(whole_data, "Triples", "null", wilcox.test)

## Applying Bonferroni correction
for(column in c(5,7,9)) {
  wx_RF_best[,column] <- p.adjust(wx_RF_best[,column], method = "bonferroni")
  wx_Tr_best[,column] <- p.adjust(wx_Tr_best[,column], method = "bonferroni")
  wx_RF_null[,column] <- p.adjust(wx_RF_null[,column], method = "bonferroni")
  wx_Tr_null[,column] <- p.adjust(wx_Tr_null[,column], method = "bonferroni")
}
```

```{r, echo = FALSE}
## Setting the table captions
captions_wx <- paste0("Non-parametric Wilcoxon test (",rep(c("Robinson-Fould distance", "Triplets distance"), 2) ," from the ",c(rep("best", 2), rep("null", 2)), " tree)" , " between the different scenarios (p-values corrected using Bonferroni-Holm correction).")

## Printing the tables
kable(wx_RF_best, row.names = FALSE, caption = captions_wx[1], digit = 3)
kable(wx_Tr_best, row.names = FALSE, caption = captions_wx[2], digit = 3)
kable(wx_RF_null, row.names = FALSE, caption = captions_wx[3], digit = 3)
kable(wx_Tr_null, row.names = FALSE, caption = captions_wx[4], digit = 3)
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_wx <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

## Tables
make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1])
make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2])
make.xtable(wx_RF_null, caption = captions_wx[3], label = labels_wx[3])
make.xtable(wx_Tr_null, caption = captions_wx[4], label = labels_wx[4])
```