### 2018-03-26 Analysis/04-EffectCorrelation_lm.Rmd
Added pooled analysis (dodgy)

---
title: "Effect of character correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Effect of character correlation

## Loading the functions and the data

Getting the functions and the data.

```{r, eval = TRUE, message = FALSE}
library(dispRity)
library(lme4)
library(lmerTest)
source("functions.R") ; load.functions(test = FALSE)
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")

## Getting the whole data
whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)
```

```{r, echo = FALSE}
library(knitr)
library(xtable)
```

# One single model example

Here we measure the eventual correlation between a low/high character correlation and a low/high RF or Triplets score.

```{r, fig.width=6, fig.height=6, message = FALSE}
## Loading the matrices CD
matrices_cd <- get.matrix.CD(path = "../Data/Simulations/Matrices", "25t_100c", verbose = TRUE, length.out = 35, cent.tend = mean)

## Running the model
model_out <- run.lm(matrices_cd, whole_data, chain = "25t_100c", method = "Bayesian", metric = "RF", cent.tend = NULL)

## Plotting the output
par(bty = "n")
one.correlation.plot(matrices_cd, whole_data, chain = "25t_100c", method = "Bayesian", metric = "RF", model_out, cent.tend = NULL, legend = TRUE)
```

# Running all the models

## Getting all matrices CD

```{r, message = FALSE}
## Loading and calculating all the matrices CDs
# all_matrices_cd <- get.all.matrix.CD(path = "../Data/Simulations/Matrices", verbose = TRUE, length.out = 35, cent.tend = mean)
# save(all_matrices_cd, file = "../Data/matrices_cd.Rda")
load("../Data/matrices_cd.Rda")
```

# Pooled models

```{r, fig.width=12, fig.height=12, warnings = FALSE}
## Pool all the results into one list
pooled_results <- pool.matrix.cd(all_matrices_cd, length.out = 28)

# ## Running the model
model_out_bay_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", cent.tend = NULL)
model_out_bay_Tr <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "Triplets", cent.tend = NULL)
model_out_par_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "RF", cent.tend = NULL)
model_out_par_Tr <- run.lm(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", cent.tend = NULL)

## Plotting the output
par(bty = "n", mfrow = c(2,2))
one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", model_out_bay_RF, cent.tend = NULL, legend = TRUE, main = "Bayesian", xlab = "RF")
one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", model_out_bay_Tr, cent.tend = NULL, legend = FALSE, main = "Bayesian", xlab = "RF")
one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", model_out_par_RF, cent.tend = NULL, legend = FALSE, main = "Parsimony", xlab = "Triplets")
one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", model_out_par_Tr, cent.tend = NULL, legend = FALSE, main = "Parsimony", xlab = "Triplets")
```

# Individual models

## Bayesian

### Robinson-Foulds

```{r}
## Running all the models
all_models_out <- run.all.lm(all_matrices_cd, whole_data, method = "Bayesian", metric = "RF")

## Getting the table
kable(correlation.table(all_models_out), digits = 3, caption = "Linear model results for the Bayesian trees and RF metric.")
```

```{r, fig.width=12, fig.height=12}
## Plotting the results
all.correlation.plot(all_matrices_cd, whole_data, metric = "RF", method = "Bayesian", models = all_models_out, col = c("blue", "orange", "green"), rounding = 3, cent.tend = NULL)
```

### Triplets

```{r}
## Running all the models
all_models_out <- run.all.lm(all_matrices_cd, whole_data, method = "Bayesian", metric = "Triplets")

## Getting the table
kable(correlation.table(all_models_out), digits = 3, caption = "Linear model results for the Bayesian trees and Triples metric.")

```

```{r, fig.width=12, fig.height=12}
## Plotting the results
all.correlation.plot(all_matrices_cd, whole_data, metric = "Triplets", method = "Bayesian", models = all_models_out, col = c("blue", "orange", "green"), rounding = 3, cent.tend = NULL)
```

## Parsimony

### Robinson-Foulds

```{r}
## Running all the models
all_models_out <- run.all.lm(all_matrices_cd, whole_data, method = "Parsimony", metric = "RF")

## Getting the table
kable(correlation.table(all_models_out), digits = 3, caption = "Linear model results for the Parsimony trees and RF metric.")

```

```{r, fig.width=12, fig.height=12}
## Plotting the results
all.correlation.plot(all_matrices_cd, whole_data, metric = "RF", method = "Parsimony", models = all_models_out, col = c("blue", "orange", "green"), rounding = 3, cent.tend = NULL)
```

### Triplets

```{r}
## Running all the models
all_models_out <- run.all.lm(all_matrices_cd, whole_data, method = "Parsimony", metric = "Triplets")

## Getting the table
kable(correlation.table(all_models_out), digits = 3, caption = "Linear model results for the Parsimony trees and Triples metric.")

```

```{r, fig.width=12, fig.height=12}
## Plotting the results
all.correlation.plot(all_matrices_cd, whole_data, metric = "Triplets", method = "Parsimony", models = all_models_out, col = c("blue", "orange", "green"), rounding = 3, cent.tend = NULL)
```