### 2018-03-27 Writing/Tables/Full_Tab_pooledsmethods_test.tex
Final results tables

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 13:20:38 2018
\begin{table}[ht]
\centering
\begin{tabular}{lllrrr}
  \hline
comp & metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
Best & RF & bayesian:parsimony & 0.891 & 579437.500 & 0.000 \\ 
   &  & bayesian:parsimony & 0.984 & 470621.500 & 0.168 \\ 
   & Tr & bayesian:parsimony & 0.847 & 657875.500 & 0.000 \\ 
   &  & bayesian:parsimony & 0.975 & 499754.000 & 0.000 \\ 
  Null & RF & bayesian:parsimony & 0.891 & 579437.500 & 0.000 \\ 
   &  & bayesian:parsimony & 0.984 & 470621.500 & 0.168 \\ 
   & Tr & bayesian:parsimony & 0.847 & 657875.500 & 0.000 \\ 
   &  & bayesian:parsimony & 0.975 & 499754.000 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Full_Tab_pooledsmethods_test}
\end{table}