### 2018-03-27 Writing/Tables/Tab_pooledmethod.tex

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:14 2018
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrrr}
  \hline
metric & method & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
RF & bayesian & 0.319 & 0.714 & 0.828 & 0.811 & 0.934 & 1.000 \\ 
   & parsimony & 0.076 & 0.496 & 0.679 & 0.684 & 0.911 & 1.000 \\ 
  Tr & bayesian & -0.473 & 0.354 & 0.738 & 0.618 & 0.947 & 1.000 \\ 
   & parsimony & -0.436 & 0.258 & 0.601 & 0.568 & 0.949 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Summary statistics of the normalised distances to the best or null tree for the pooled methods.} 
\label{Tab_pooledmethod}
\end{table}