### 2018-03-27 Writing/Tables/Tab_pooledsmethods_test.tex

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:14 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & bayesian:parsimony & 0.891 & 579437.500 & \textbf{0} \\ 
  Tr & bayesian:parsimony & \textbf{0.984} & 470621.500 & 0.084 \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Tab_pooledsmethods_test}
\end{table}