### 2018-04-10 Writing/CoverLetter.tex
Rewrote the cover letter

\documentclass[11pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
\usepackage[osf]{mathpazo}
\signature{Thomas Guillerme \\ Martin Brazeau}
\address{The University of Queensland \\School of Biological Sciences \\St Lucia QLD 4067, Australia \\guillert@tcd.ie}
\longindentation=0pt
\begin{document}

\begin{letter}{}
\opening{Dear Editors,}

Despite being at the base of almost every endeavour in modern Biology, the assumption made in tree inference that phylogenetic data is independent is probably mostly erroneous.
This assumption was stated since the advent of Cladistics in the 1950s but is nowadays mainly ignored in phylogenetic endeavour.
In fact, this effect can be neglected in molecular datasets because the huge amount of available data is thought to temper down the correlation effect.
However, in morphological datasets, where the amount of characters is much smaller, character correlation can lead to significant biases especially by favouring one topology from another due to a specific choice of characters (whether this is done consciously by researchers or not).
With the ``resurgence'' of morphological data phylogenetic analysis (namely with the Total Evidence tip-dating method, Ronquist \textit{et al.}, 2012, Sys. Biol), it is thus important to acknowledge limitations and caveats that come with the use of such data, namely how correlation patterns in morphological datasets can effect tree inference.

In our research article, entitled ``Influence of different modes of morphological character correlation on tree inference'', we propose the first - to our knowledge - simulation based test of the effect of character correlation on tree topology using a thorough simulation framework ($>$ 100 CPU years worth).
Using a novel metric to measure direct phylogenetic correlation between characters, we simulated discrete morphological datasets that we then modified to increase or decrease character correlation.
We then measured the resulting topological differences (from both maximum parsimony and Bayesian trees) to measure the effect of character correlation in the light of our known simulated parameters.

Our results suggest that datasets that have higher levels of internal correlation negatively impact our ability to recover the correct tree topology.
This is too be contrasted with datasets that have low levels of correlation (i.e. more noisy) which can still recover the correct topology, depending on the amount of data and the inference method used.
These results, although non-surprising to researchers in the field, are the first - to our knowledge - to actually quantify the effect of character correlation and provides practical data and reusable code allowing researcher to effectively measure the limitations of their datasets.
Furthermore, we suggest that using our proposed new metric, it would be able to ``bias'' bootstrap pseudo-replications in phylogenetic analysis by favouring character with high or low correlation, depending on the purpose of the phylogenetic search.

We look forward to hearing from you soon,

\closing{Yours sincerely,}

\end{letter}
\end{document}

%Editors:
%Sam Price, Daniele Silvestro

%Reviewers
%Graeme Lloyd, Dave Bapst, Emma Sherratt