### 2018-06-04 Writing/SystematicBiology/Tab_pooledstaxa_test.tex

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:13 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & t25:t75 & \textbf{0.976} & 218421.000 & \textbf{0.012} \\ 
   & t25:t150 & \textbf{0.988} & 220529.000 & \textbf{0.004} \\ 
   & t75:t150 & \textbf{0.99} & 201037.000 & 1 \\ 
  Tr & t25:t75 & \textbf{0.976} & 233282.000 & \textbf{0} \\ 
   & t25:t150 & \textbf{0.978} & 227288.000 & \textbf{0} \\ 
   & t75:t150 & \textbf{0.992} & 194201.000 & 1 \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled number of taxa. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledstaxa_test}
\end{table}