### 2018-06-20 Analysis/04-CheckMatrixCorrelation.Rmd
Added statistics table

---
title: "Check matrix correlations"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Checking the number of character replaced in each matrix

```{r}
library(dispRity)
library(ape)
library(knitr)
## Setting the path
PATH <- "../Data/Simulations/Matrices/"
```

The following function counts, _a posteriori_ the number of characters that where replaced in each matrix (by comparing them to the normal matrix)

```{r}
## Getting the proportion of replaced characters
get.proportion.replace <- function(matrix, path) {
    ## Getting one matrix in a comparable format
    get.one.matrix <- function(path, matrix, type) {
        return(unlist(apply(do.call(rbind,
                                    read.nexus.data(paste0(path, matrix, "_", type, ".nex"))
                            ), 2, list), recursive = FALSE))
    }

    ## Getting all the matrices
    norm <- get.one.matrix(path, matrix, "norm")
    maxi <- get.one.matrix(path, matrix, "maxi")
    mini <- get.one.matrix(path, matrix, "mini")
    rand <- get.one.matrix(path, matrix, "rand")

    ## Comparing all the matrices
    n <- length(norm)
    maxi <- length(which(!mapply(function(X, Y) all(X == Y), norm, maxi)))/n
    mini <- length(which(!mapply(function(X, Y) all(X == Y), norm, mini)))/n
    rand <- length(which(!mapply(function(X, Y) all(X == Y), norm, rand)))/n

    return(c("maxi" = maxi, "mini" = mini, "rand" = rand))
}
```

We can then apply it to the list of matrices

```{r}
## Creating the placeholder list for the results
replacements <- list(list(), list(), list())
## getting the values to loop through
taxa_val <- c(25, 75, 150)
character_val <- c(100, 350, 1000)
for(taxa in 1:3) {
    for(character in 1:3) {
        ## Get the taxa names
        names <- unique(gsub("_[a-z]*.nex", "",
          list.files(PATH, pattern = paste0(taxa_val[taxa], "t_", character_val[character], "c_"))))
        ## Get the proportions of replaced characters
        replacements[[taxa]][[character]] <- lapply(as.list(names), get.proportion.replace, PATH)
    }
}

## Combining the results
fun.maxi <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[1]]))))
}
fun.mini <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[2]]))))
}
fun.rand <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[3]]))))
}
results_maxi <- lapply(replacements, lapply, fun.maxi)
results_mini <- lapply(replacements, lapply, fun.mini)
results_rand <- lapply(replacements, lapply, fun.rand)

## Naming the lists
c("25t", "75t", "150t") -> names(results_maxi) -> names(results_mini) -> names(results_rand)
c("100c", "350c", "1000c") -> names(results_maxi[[1]]) ->
                              names(results_maxi[[2]]) ->
                              names(results_maxi[[3]]) ->
                              names(results_mini[[1]]) ->
                              names(results_mini[[2]]) ->
                              names(results_mini[[3]]) ->
                              names(results_rand[[1]]) ->
                              names(results_rand[[2]]) ->
                              names(results_rand[[3]])
```

We can then represent the results using box plots:

```{r, fig.height = 18, fig.width = 6}
## Plotting the different proportion results
plot.proportions <- function(data, ...) {
    ## Default colours
    cols <- gray.colors(3)
    ## Boxes
    boxplot(xlab = "", ylab = "Proportion of replacement", xaxt = "n", ylim = c(0, 1), col = cols,
            c(data[[1]], data[[2]], data[[3]]), ...)
    legend("topleft", legend = names(data), pch = 19, col = cols, bty = "n")
    ## X axis
    axis(1, 1:9, labels = FALSE, tick = FALSE)
    axis(1, c(2, 5, 8), tick = FALSE, labels = names(data[[1]]))
    ## Lines
    abline(v = 3.5) ; abline(v = 6.5)
}

par(mfrow = c(3,1))
plot.proportions(results_maxi, main = "Maximised differences")
plot.proportions(results_mini, main = "Minimised differences")
plot.proportions(results_rand, main = "Randomised differences")

```

<!-- Figure version -->

```{r, fig.height = 18, fig.width = 6, eval = TRUE, echo = FALSE}
pdf("../Writing/Figures/Proportion_replaced_maxi.pdf", width = 6, height = 6)
plot.proportions(results_maxi, main = "Maximised differences")
pdf("../Writing/Figures/Proportion_replaced_mini.pdf", width = 6, height = 6)
plot.proportions(results_mini, main = "Minimised differences")
pdf("../Writing/Figures/Proportion_replaced_rand.pdf", width = 6, height = 6)
plot.proportions(results_rand, main = "Randomised differences")
```

## Testing the differences in proportions

We can then test the differences between categories.
First we'll try an aov:

```{r}
## Data frame converter
convert.df <- function(results) {
    convert.sub.df <- function(list, names) {
        df <- data.frame(unlist(list), as.factor(rep(names(list), each = length(list[[1]]))))
        colnames(df) <- names
        rownames(df) <- NULL
        return(df)
    }
    df_list <- lapply(results, convert.sub.df, names = c("proportion", "characters"))
    df_list <- cbind(do.call(rbind, df_list),
                     "taxa" = as.factor(rep(names(results), each = nrow(df_list[[1]]))))
    rownames(df_list) <- NULL
    return(df_list)
}

## Generating the nested tables
table_maxi <- convert.df(results_maxi)
table_mini <- convert.df(results_mini)
table_rand <- convert.df(results_rand)

## Combining the tables
all_results <- cbind(rbind(table_maxi, table_mini, table_rand),
                    "scenario" = as.factor(rep(c("maxi", "mini", "rand"), each = nrow(table_maxi))))

## Run the linear model for testing assumptions
model <- lm(proportion ~ scenario, data = all_results)
(residuals_normality <- shapiro.test(residuals(model)))
(data_homostedasticity <- bartlett.test(proportion ~ scenario, data = all_results))

## Normality assumption fails
```

However, the assumptions are violated so we need to go non-parametric (Wilcox test and Bhattacharyya Coefficient)

```{r}
## Pairwise tests wrapper
pair.test <- function(data, factor, test) {
    ## Separating the data according to factors
    factor <- which(colnames(data) == factor)
    splited_data <- split(data, data[, factor])

    ## Combination of tests
    combinations <- combn(1:length(splited_data), 2)

    ## Applying the tests
    tests <- apply(combinations, 2, function(X, test, data)
                 return(test(data[[X[1]]][,1], data[[X[2]]][,1])), test = test, data = splited_data)
    names_test <- apply(combinations, 2, function(X, data)
                   return(paste0(names(data)[X[1]], " - ", names(data)[X[2]])), data = splited_data)

    names(tests) <- names_test
    return(tests)
}

## Running the wilcox tests for each pooled factor
wilcox_scen <- pair.test(all_results, factor = "scenario", wilcox.test)
wilcox_taxa <- pair.test(all_results, factor = "taxa", wilcox.test)
wilcox_char <- pair.test(all_results, factor = "characters", wilcox.test)

## Extracting the statistic and the p_value only
wilcox_scen_w <- unlist(lapply(wilcox_scen, function(X) return(unname(X$statistic))))
wilcox_taxa_w <- unlist(lapply(wilcox_taxa, function(X) return(unname(X$statistic))))
wilcox_char_w <- unlist(lapply(wilcox_char, function(X) return(unname(X$statistic))))
wilcox_scen_p <- p.adjust(unlist(lapply(wilcox_scen, function(X) return(X$p.value))), "bonferroni")
wilcox_taxa_p <- p.adjust(unlist(lapply(wilcox_taxa, function(X) return(X$p.value))), "bonferroni")
wilcox_char_p <- p.adjust(unlist(lapply(wilcox_char, function(X) return(X$p.value))), "bonferroni")

## Running the Bhattacharrya tests for each pooled factor
bhattc_scen <- pair.test(all_results, factor = "scenario", bhatt.coeff)
bhattc_taxa <- pair.test(all_results, factor = "taxa", bhatt.coeff)
bhattc_char <- pair.test(all_results, factor = "characters", bhatt.coeff)

## Combine the results
combined_results <- as.data.frame(rbind(
        cbind("statistic" = wilcox_scen_w, "p.value" = wilcox_scen_p, "bhatt.coeff" = bhattc_scen),
        cbind("statistic" = wilcox_taxa_w, "p.value" = wilcox_taxa_p, "bhatt.coeff" = bhattc_taxa),
        cbind("statistic" = wilcox_char_w, "p.value" = wilcox_char_p, "bhatt.coeff" = bhattc_char)))
```

And here are the results:

```{r, echo = FALSE}
caption <- paste0("Pairwise Wilcoxon test and Bhattacharrya Coefficients for the difference in",
  " proportion of character replacement for the different scenarios, number of taxa and characters.")
kable(combined_results, digits = 3, caption = caption)
```

```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
library(xtable)
source("../Functions/make.xtable.R")
make.xtable(combined_results, caption = caption, label = "Tab_proportion_replacement",
            path = "../Writing/Tables/", include.rownames = TRUE)
```