### 2019-04-16 Writing/Tables/Tab_pooledmethod.tex

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 16 16:08:54 2019
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrrr}
  \hline
metric & method & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
RF & bayesian & 0.385 & 0.753 & 0.851 & 0.824 & 0.924 & 1.000 \\ 
   & parsimony & 0.371 & 0.642 & 0.739 & 0.742 & 0.865 & 1.000 \\ 
  Tr & bayesian & -0.196 & 0.560 & 0.802 & 0.717 & 0.945 & 1.000 \\ 
   & parsimony & -0.155 & 0.525 & 0.750 & 0.693 & 0.900 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Summary statistics of the normalised distances to the best or null tree for the pooled methods.} 
\label{Tab_pooledmethod}
\end{table}