### 2019-04-16 Writing/Tables/Tab_pooledscharacters_test.tex

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 16 16:07:59 2019
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100:c350 & \textbf{0.965} & 48263.500 & \textbf{0.000} \\ 
   & c100:c1000 & 0.914 & 37714.500 & \textbf{0.000} \\ 
   & c350:c1000 & \textbf{0.976} & 53138.000 & \textbf{0.000} \\ 
  Tr & c100:c350 & 0.929 & 40253.000 & \textbf{0.000} \\ 
   & c100:c1000 & 0.826 & 27923.000 & \textbf{0.000} \\ 
   & c350:c1000 & \textbf{0.957} & 48164.500 & \textbf{0.000} \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled number of characters. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Tab_pooledscharacters_test}
\end{table}