### 2019-04-16 Writing/Tables/Tab_pooledsmethods_test.tex

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 16 16:09:14 2019
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & bayesian:parsimony & 0.939 & 192659.000 & \textbf{0.000} \\ 
  Tr & bayesian:parsimony & \textbf{0.986} & 157677.500 & \textbf{0.041} \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Tab_pooledsmethods_test}
\end{table}