### 2019-04-17 Analysis/04-CheckMatrixCorrelation.Rmd
Updated character correlations

---
title: "Check matrix correlations"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Checking the number of character replaced in each matrix and the change in character correlation

```{r}
library(dispRity)
library(ape)
library(knitr)
## Setting the path
PATH <- "../Data/Simulations/"
```

The following function counts the effective change in number of characters for each matrix: the average character correlation change, their distributions, the number of changed characters, and their character correlation change.

## Load the data

We can use the following functions to load the already estimated parameters (`read.matrix.param`) during the simulations or directly the matrices (`read.matrix`) to specifically double check the parameters (not in this document).

```{r}
## Loading a matrix
read.matrix <- function(chain_name, path) {
    ## Read the four matrices
    norm <- do.call(rbind, read.nexus.data(paste0(path, chain_name, "_norm.nex")))
    maxi <- do.call(rbind, read.nexus.data(paste0(path, chain_name, "_maxi.nex")))
    mini <- do.call(rbind, read.nexus.data(paste0(path, chain_name, "_mini.nex")))
    rand <- do.call(rbind, read.nexus.data(paste0(path, chain_name, "_rand.nex")))
    return(list(norm = norm, maxi = maxi, mini = mini, rand = rand))
}

## Loading a matrix parameter
read.matrix.param <- function(chain_name, path) {
    ## Reading the parameters
    maxi <- read.csv(paste0(path, chain_name, "_changes_maxi.csv"), row.names = 1)
    mini <- read.csv(paste0(path, chain_name, "_changes_mini.csv"), row.names = 1)
    rand <- read.csv(paste0(path, chain_name, "_changes_rand.csv"), row.names = 1)

    ## Splitting the parameters
    char_diff_change <- rbind(maxi = maxi[1,], mini = mini[1,], rand = rand[1,])
    colnames(char_diff_change) <- c("original", "modified")

    ## Getting the list of changes
    characters_changes <- list(maxi = maxi[-1,], mini = mini[-1,], rand = rand[-1,])

    ## Return both
    return(list(differences = char_diff_change, changes = characters_changes))
}
```

We can load all the parameters arranged into a list of list of scenarios (e.g. 25 taxa and 100 characters, etc.):

```{r, echo = FALSE}
## Creating the placeholder list for the results
parameters <- list(list(), list(), list())
## getting the values to loop through
taxa_val <- c(25, 75, 150)
character_val <- c(100, 350, 1000)
## Set the path
param_path <- paste0(PATH, "Parameters/")

## Loading the parameters
for(taxa in 1:3) {
    for(character in 1:3) {
        ## Get the chain names
        chain_names <- list.files(path = param_path, pattern = paste0(taxa_val[taxa], "t_", character_val[character], "c_"))
        ## Remove the suffix
        chain_names <- unique(gsub("_changes_\\S+.csv", "", chain_names, perl = TRUE))

        ## Get the matrix change parameters
        parameters[[taxa]][[character]] <- lapply(as.list(chain_names), read.matrix.param, path = param_path)
    }
    names(parameters[[taxa]]) <- paste0(character_val, "c")
}
names(parameters) <- paste0(taxa_val, "t")
```

## Average character difference change

Getting the difference between the character difference in the normal matrix and the one in the modified ones:

```{r}
## Get character difference change value
get.char.diff.change <- function(chain) {
    return(chain$differences)
}
parameters_difference <- lapply(parameters, lapply, lapply, get.char.diff.change)

## Get the differences
parameters_diff <- lapply(parameters_difference, lapply, lapply, function(X) apply(X, 1, diff))
```

Plotting the results

```{r, fig.height = 18, fig.width = 6}
## Extract the results for each type of scenarios
extract.results <- function(list, what) {
    return(unname(unlist(lapply(list, function(X) return(X[what])))))
}
differences_maxi <- lapply(parameters_diff, lapply, extract.results, what = "maxi")
differences_mini <- lapply(parameters_diff, lapply, extract.results, what = "mini")
differences_rand <- lapply(parameters_diff, lapply, extract.results, what = "rand")

## overal_range (for plotting)
overal_range <- range(c(unlist(differences_maxi), unlist(differences_mini), unlist(differences_rand)))

## Plotting the different proportion results
plot.proportions <- function(data, ylim = c(-1, 1), xlim = c(0.8, 9.2), leg.pos, ylab, ...) {
    ## Default colours
    cols <- gray.colors(length(data))
    ## Boxes
    boxplot(NULL, xlab = "", ylab = ylab, xaxt = "n", yaxt = "n",  ylim = ylim, xlim = xlim, ...)
    legend(leg.pos, legend = names(data), pch = 19, col = cols, bty = "n")
    ## X axis
    axis(1, 1:9, labels = FALSE, tick = FALSE)
    axis(1, c(2, 5, 8), tick = FALSE, labels = names(data[[1]]))
    ## Lines
    abline(v = 3.5) ; abline(v = 6.5) ; abline(h = 0, lty = 2)

    ## Data
    boxplot(c(data[[1]], data[[2]], data[[3]]), col = cols, add = TRUE, xaxt = "n")
}
par(mfrow = c(3,1))
plot.proportions(differences_maxi, ylim = overal_range, main = "Maximised", leg.pos = "bottomleft",
    ylab = "Character differences change")
plot.proportions(differences_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
    ylab = "Character differences change")
plot.proportions(differences_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft",
    ylab = "Character differences change")
```

<!-- Figure version -->

```{r, fig.height = 18, fig.width = 6, eval = TRUE, echo = FALSE}
pdf("../Writing/Figures/Difference_change_maxi.pdf", width = 6, height = 6)
plot.proportions(differences_maxi, ylim = overal_range, main = "Maximised", leg.pos = "bottomleft",
    ylab = "Character differences change")
pdf("../Writing/Figures/Difference_change_mini.pdf", width = 6, height = 6)
plot.proportions(differences_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
    ylab = "Character differences change")
pdf("../Writing/Figures/Difference_change_rand.pdf", width = 6, height = 6)
plot.proportions(differences_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft",
    ylab = "Character differences change")
```

### Testing the differences in character difference

We can test the differences in character differences using the combined Wilcoxon test and Bhattacharyya coefficients.

First we need to convert it into a usable table format:

```{r}
## Data frame converter
convert.df <- function(results) {
    convert.sub.df <- function(list, names) {
        df <- data.frame(unlist(list), as.factor(rep(names(list), each = length(list[[1]]))))
        colnames(df) <- names
        rownames(df) <- NULL
        return(df)
    }
    df_list <- lapply(results, convert.sub.df, names = c("proportion", "characters"))
    df_list <- cbind(do.call(rbind, df_list),
                     "taxa" = as.factor(rep(names(results), each = nrow(df_list[[1]]))))
    rownames(df_list) <- NULL
    return(df_list)
}

## Generating the nested tables
table_maxi <- convert.df(differences_maxi)
table_mini <- convert.df(differences_mini)
table_rand <- convert.df(differences_rand)

## Combining the tables
all_results <- cbind(rbind(table_maxi, table_mini, table_rand),
                    "scenario" = as.factor(rep(c("maxi", "mini", "rand"), each = nrow(table_maxi))))
```

```{r}
## Pairwise tests wrapper
pair.test <- function(data, factor, test) {
    ## Separating the data according to factors
    factor <- which(colnames(data) == factor)
    splited_data <- split(data, data[, factor])

    ## Combination of tests
    combinations <- combn(1:length(splited_data), 2)

    ## Applying the tests
    tests <- apply(combinations, 2, function(X, test, data)
                 return(test(data[[X[1]]][,1], data[[X[2]]][,1])), test = test, data = splited_data)
    names_test <- apply(combinations, 2, function(X, data)
                   return(paste0(names(data)[X[1]], " - ", names(data)[X[2]])), data = splited_data)

    names(tests) <- names_test
    return(tests)
}

## Running the wilcox tests for each pooled factor
wilcox_scen <- pair.test(all_results, factor = "scenario", wilcox.test)
wilcox_taxa <- pair.test(all_results, factor = "taxa", wilcox.test)
wilcox_char <- pair.test(all_results, factor = "characters", wilcox.test)

## Extracting the statistic and the p_value only
wilcox_scen_w <- unlist(lapply(wilcox_scen, function(X) return(unname(X$statistic))))
wilcox_taxa_w <- unlist(lapply(wilcox_taxa, function(X) return(unname(X$statistic))))
wilcox_char_w <- unlist(lapply(wilcox_char, function(X) return(unname(X$statistic))))
wilcox_scen_p <- p.adjust(unlist(lapply(wilcox_scen, function(X) return(X$p.value))), "bonferroni")
wilcox_taxa_p <- p.adjust(unlist(lapply(wilcox_taxa, function(X) return(X$p.value))), "bonferroni")
wilcox_char_p <- p.adjust(unlist(lapply(wilcox_char, function(X) return(X$p.value))), "bonferroni")

## Running the Bhattacharrya tests for each pooled factor
bhattc_scen <- pair.test(all_results, factor = "scenario", bhatt.coeff)
bhattc_taxa <- pair.test(all_results, factor = "taxa", bhatt.coeff)
bhattc_char <- pair.test(all_results, factor = "characters", bhatt.coeff)

## Combine the results
combined_results <- as.data.frame(rbind(
        cbind("statistic" = wilcox_scen_w, "p.value" = wilcox_scen_p, "bhatt.coeff" = bhattc_scen),
        cbind("statistic" = wilcox_taxa_w, "p.value" = wilcox_taxa_p, "bhatt.coeff" = bhattc_taxa),
        cbind("statistic" = wilcox_char_w, "p.value" = wilcox_char_p, "bhatt.coeff" = bhattc_char)))
```

And here are the results:

```{r, echo = FALSE}
caption <- paste0("Pairwise Wilcoxon test and Bhattacharrya Coefficients for the differences in",
  "character difference for the different scenarios, number of taxa and characters.")
kable(combined_results, digits = 3, caption = caption)
```

```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
library(xtable)
source("../Functions/make.xtable.R")
make.xtable(combined_results, caption = caption, label = "Tab_difference_change",
            path = "../Writing/Tables/", include.rownames = TRUE)
```

## Number of characters duplicated

We can also look at the number of characters that are duplicated in each

```{r}
## Get character difference change value
get.char.replace <- function(chain) {
    return(lapply(chain$changes, function(X) return(length(unique(X[,2])))))
}
## Get the number of duplicated characters
parameters_duplicates <- lapply(parameters, lapply, lapply, get.char.replace)
```

Plotting the results

```{r, fig.height = 18, fig.width = 6}
## Make the characters proportional
make.prop <- function(list) {
    return(mapply(function(x,y) return(x/y), list, as.list(c(100, 350, 1000)),
                 SIMPLIFY = FALSE))
}

## Extract the results for each type of scenarios
duplicates_maxi <- lapply(lapply(parameters_duplicates, lapply, extract.results, what = "maxi"),
                        make.prop)
duplicates_mini <- lapply(lapply(parameters_duplicates, lapply, extract.results, what = "mini"),
                        make.prop)
duplicates_rand <- lapply(lapply(parameters_duplicates, lapply, extract.results, what = "rand"),
                        make.prop)

## overal_range (for plotting)
overal_range <- c(0,1)

## Plotting the different proportion results
par(mfrow = c(3,1))
plot.proportions(duplicates_maxi, ylim = overal_range, main = "Maximised", leg.pos = "topleft",
    ylab = "Proportion of duplicated characters")
plot.proportions(duplicates_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
    ylab = "Proportion of duplicated characters")
plot.proportions(duplicates_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft",
    ylab = "Proportion of duplicated characters")
```

<!-- Figure version -->

```{r, fig.height = 18, fig.width = 6, eval = TRUE, echo = FALSE}
pdf("../Writing/Figures/Proportion_duplicated_maxi.pdf", width = 6, height = 6)
plot.proportions(duplicates_maxi, ylim = overal_range, main = "Maximised", leg.pos = "topleft",
    ylab = "Proportion of duplicated characters")
pdf("../Writing/Figures/Proportion_duplicated_mini.pdf", width = 6, height = 6)
plot.proportions(duplicates_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
    ylab = "Proportion of duplicated characters")
pdf("../Writing/Figures/Proportion_duplicated_rand.pdf", width = 6, height = 6)
plot.proportions(duplicates_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft",
    ylab = "Proportion of duplicated characters")
```

### Testing the differences in character difference

We can test the differences in character differences using the combined Wilcoxon test and Bhattacharyya coefficients.

First we need to convert it into a usable table format:

```{r}
## Generating the nested tables
table_maxi <- convert.df(duplicates_maxi)
table_mini <- convert.df(duplicates_mini)
table_rand <- convert.df(duplicates_rand)

## Combining the tables
all_results <- cbind(rbind(table_maxi, table_mini, table_rand),
                    "scenario" = as.factor(rep(c("maxi", "mini", "rand"), each = nrow(table_maxi))))
```

```{r}
## Running the wilcox tests for each pooled factor
wilcox_scen <- pair.test(all_results, factor = "scenario", wilcox.test)
wilcox_taxa <- pair.test(all_results, factor = "taxa", wilcox.test)
wilcox_char <- pair.test(all_results, factor = "characters", wilcox.test)

## Extracting the statistic and the p_value only
wilcox_scen_w <- unlist(lapply(wilcox_scen, function(X) return(unname(X$statistic))))
wilcox_taxa_w <- unlist(lapply(wilcox_taxa, function(X) return(unname(X$statistic))))
wilcox_char_w <- unlist(lapply(wilcox_char, function(X) return(unname(X$statistic))))
wilcox_scen_p <- p.adjust(unlist(lapply(wilcox_scen, function(X) return(X$p.value))), "bonferroni")
wilcox_taxa_p <- p.adjust(unlist(lapply(wilcox_taxa, function(X) return(X$p.value))), "bonferroni")
wilcox_char_p <- p.adjust(unlist(lapply(wilcox_char, function(X) return(X$p.value))), "bonferroni")

## Running the Bhattacharrya tests for each pooled factor
bhattc_scen <- pair.test(all_results, factor = "scenario", bhatt.coeff)
bhattc_taxa <- pair.test(all_results, factor = "taxa", bhatt.coeff)
bhattc_char <- pair.test(all_results, factor = "characters", bhatt.coeff)

## Combine the results
combined_results <- as.data.frame(rbind(
        cbind("statistic" = wilcox_scen_w, "p.value" = wilcox_scen_p, "bhatt.coeff" = bhattc_scen),
        cbind("statistic" = wilcox_taxa_w, "p.value" = wilcox_taxa_p, "bhatt.coeff" = bhattc_taxa),
        cbind("statistic" = wilcox_char_w, "p.value" = wilcox_char_p, "bhatt.coeff" = bhattc_char)))
```

And here are the results:

```{r, echo = FALSE}
caption <- paste0("Pairwise Wilcoxon test and Bhattacharrya Coefficients for the difference in",
  " proportion of character duplicated for the different scenarios, number of taxa and characters.")
kable(combined_results, digits = 3, caption = caption)
```

```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
library(xtable)
source("../Functions/make.xtable.R")
make.xtable(combined_results, caption = caption, label = "Tab_proportion_duplicated",
            path = "../Writing/Tables/", include.rownames = TRUE)
```