### 2019-04-23 Writing/Tables/Full_Tab_Bayesian_WXRFbest.tex
added combined results test tables

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
bayesian & 25 & 100 & 136.500 & 1.000 & 167.000 & 1.000 & 219.000 & 1.000 \\ 
   &  & 350 & 123.500 & 0.707 & 148.000 & 1.000 & 216.000 & 1.000 \\ 
   &  & 1000 & 51.500 & 0.001 & 81.000 & 0.024 & 218.000 & 1.000 \\ 
   & 75 & 100 & 5.500 & 0.000 & 51.500 & 0.001 & 381.000 & 0.000 \\ 
   &  & 350 & 20.000 & 0.000 & 52.500 & 0.001 & 301.500 & 0.112 \\ 
   &  & 1000 & 84.000 & 0.032 & 118.000 & 0.494 & 296.000 & 0.174 \\ 
   & 150 & 100 & 25.500 & 0.000 & 108.000 & 0.239 & 340.000 & 0.003 \\ 
   &  & 350 & 69.500 & 0.008 & 121.500 & 0.625 & 327.500 & 0.011 \\ 
   &  & 1000 & 105.500 & 0.198 & 122.500 & 0.669 & 252.500 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Robinson-Fould distance from the best tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Bayesian_WXRFbest}
\end{table}