### 2019-04-23 Writing/Tables/Full_Tab_Bayesian_WXTrnull.tex

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
bayesian & 25 & 100 & 244.000 & 1.000 & 300.500 & 0.123 & 246.500 & 1.000 \\ 
   &  & 350 & 257.500 & 1.000 & 254.000 & 1.000 & 205.500 & 1.000 \\ 
   &  & 1000 & 316.000 & 0.032 & 243.500 & 1.000 & 130.000 & 1.000 \\ 
   & 75 & 100 & 255.000 & 1.000 & 223.000 & 1.000 & 168.000 & 1.000 \\ 
   &  & 350 & 278.000 & 0.648 & 260.000 & 1.000 & 165.500 & 1.000 \\ 
   &  & 1000 & 238.000 & 1.000 & 258.000 & 1.000 & 212.000 & 1.000 \\ 
   & 150 & 100 & 240.000 & 1.000 & 188.000 & 1.000 & 147.000 & 1.000 \\ 
   &  & 350 & 229.000 & 1.000 & 229.000 & 1.000 & 198.000 & 1.000 \\ 
   &  & 1000 & 286.500 & 0.360 & 252.000 & 1.000 & 176.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Triplets distance from the null tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Bayesian_WXTrnull}
\end{table}