### 2019-04-23 Writing/Tables/Full_Tab_Parsimony_WXRFbest.tex

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
parsimony & 25 & 100 & 159.000 & 1.000 & 215.500 & 1.000 & 266.000 & 1.000 \\ 
   &  & 350 & 127.500 & 0.913 & 153.000 & 1.000 & 215.000 & 1.000 \\ 
   &  & 1000 & 123.500 & 0.698 & 131.000 & 1.000 & 207.000 & 1.000 \\ 
   & 75 & 100 & 44.000 & 0.000 & 117.000 & 0.460 & 313.000 & 0.042 \\ 
   &  & 350 & 100.500 & 0.133 & 123.500 & 0.714 & 247.000 & 1.000 \\ 
   &  & 1000 & 86.500 & 0.040 & 112.000 & 0.322 & 275.500 & 0.762 \\ 
   & 150 & 100 & 111.000 & 0.299 & 188.500 & 1.000 & 274.500 & 0.814 \\ 
   &  & 350 & 95.000 & 0.084 & 165.000 & 1.000 & 301.000 & 0.117 \\ 
   &  & 1000 & 91.000 & 0.060 & 122.000 & 0.648 & 240.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Robinson-Fould distance from the best tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Parsimony_WXRFbest}
\end{table}