### 2019-04-23 Writing/Tables/Full_Tab_Parsimony_WXRFnull.tex

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
parsimony & 25 & 100 & 281.000 & 0.521 & 259.500 & 1.000 & 178.500 & 1.000 \\ 
   &  & 350 & 254.500 & 1.000 & 244.500 & 1.000 & 190.000 & 1.000 \\ 
   &  & 1000 & 280.500 & 0.532 & 225.500 & 1.000 & 149.500 & 1.000 \\ 
   & 75 & 100 & 322.500 & 0.017 & 232.500 & 1.000 & 75.500 & 0.014 \\ 
   &  & 350 & 320.500 & 0.021 & 267.500 & 1.000 & 115.500 & 0.414 \\ 
   &  & 1000 & 320.000 & 0.022 & 250.500 & 1.000 & 111.000 & 0.297 \\ 
   & 150 & 100 & 265.500 & 1.000 & 260.500 & 1.000 & 182.500 & 1.000 \\ 
   &  & 350 & 301.000 & 0.117 & 251.500 & 1.000 & 136.000 & 1.000 \\ 
   &  & 1000 & 274.500 & 0.815 & 258.000 & 1.000 & 157.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Robinson-Fould distance from the null tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Parsimony_WXRFnull}
\end{table}