### 2019-04-23 Writing/Tables/Full_Tab_Parsimony_WXTrnull.tex

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
parsimony & 25 & 100 & 284.500 & 0.415 & 285.500 & 0.386 & 193.500 & 1.000 \\ 
   &  & 350 & 223.000 & 1.000 & 253.000 & 1.000 & 244.000 & 1.000 \\ 
   &  & 1000 & 262.000 & 1.000 & 231.000 & 1.000 & 179.000 & 1.000 \\ 
   & 75 & 100 & 245.000 & 1.000 & 248.000 & 1.000 & 222.000 & 1.000 \\ 
   &  & 350 & 246.000 & 1.000 & 253.000 & 1.000 & 208.000 & 1.000 \\ 
   &  & 1000 & 242.500 & 1.000 & 257.000 & 1.000 & 206.000 & 1.000 \\ 
   & 150 & 100 & 240.000 & 1.000 & 236.000 & 1.000 & 177.000 & 1.000 \\ 
   &  & 350 & 213.000 & 1.000 & 198.000 & 1.000 & 203.000 & 1.000 \\ 
   &  & 1000 & 261.000 & 1.000 & 274.000 & 0.827 & 222.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Triplets distance from the null tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Parsimony_WXTrnull}
\end{table}