### 2019-05-31 Writing/Tables/Full_Tab_pooledsmethods_test.tex
Updated tables

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Fri May 31 13:31:50 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrr}
  \hline
comp & metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
Best & RF & bayesian vs. parsimony & 0.939 & 192659.000 & 0.000 \\ 
   & Tr & bayesian vs. parsimony & 0.986 & 157677.500 & 0.082 \\ 
  Null & RF & bayesian vs. parsimony & 0.944 & 193142.000 & 0.000 \\ 
   & Tr & bayesian vs. parsimony & 0.975 & 161621.000 & 0.008 \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Full_Tab_pooledsmethods_test}
\end{table}