### 2019-10-15 Writing/Palaeontology/Tab_pooledsmethods_test.tex

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Fri May 31 14:42:25 2019
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & bayesian vs. parsimony & 0.939 & 192659.000 & \textbf{0.000} \\ 
  Tr & bayesian vs. parsimony & \textbf{0.986} & 157677.500 & \textbf{0.041} \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the two different inference methods: bayesian or maximum parsimony. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledsmethods_test}
\end{table}