### 2019-10-15 Writing/Palaeontology/Tab_pooledstaxa_test.tex

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Fri May 31 14:42:25 2019
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & t25 vs. t75 & 0.949 & 79945.000 & \textbf{0.000} \\ 
   & t25 vs. t150 & \textbf{0.954} & 75906.000 & \textbf{0.000} \\ 
   & t75 vs. t150 & \textbf{0.984} & 60328.000 & 0.654 \\ 
  Tr & t25 vs. t75 & \textbf{0.960} & 72814.000 & \textbf{0.024} \\ 
   & t25 vs. t150 & \textbf{0.961} & 72716.000 & \textbf{0.027} \\ 
   & t75 vs. t150 & \textbf{0.980} & 64637.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of taxa: 25 (t25), 75 (t75) and 150 (t150). Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledstaxa_test}
\end{table}