### 2022-03-24 Writing/Tables/Full_Tab_Bayesian_WXRFnull.tex

% latex table generated in R 4.1.3 by xtable 1.8-4 package
% Thu Mar 24 14:25:30 2022
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & minimised.maximised.W & minimised.maximised.p & minimised.randomised.W & minimised.randomised.p & maximised.randomised.W & maximised.randomised.p \\ 
  \hline
bayesian & 25 & 100 & 315.500 & 0.033 & 310.500 & 0.051 & 183.500 & 1.000 \\ 
   &  & 350 & 301.000 & 0.115 & 248.000 & 1.000 & 135.500 & 1.000 \\ 
   &  & 1000 & 345.000 & 0.002 & 228.000 & 1.000 & 65.500 & 0.005 \\ 
   & 75 & 100 & 353.000 & 0.001 & 252.500 & 1.000 & 59.500 & 0.003 \\ 
   &  & 350 & 355.000 & 0.001 & 258.000 & 1.000 & 64.500 & 0.005 \\ 
   &  & 1000 & 293.500 & 0.213 & 209.500 & 1.000 & 113.500 & 0.359 \\ 
   & 150 & 100 & 310.500 & 0.053 & 226.000 & 1.000 & 98.000 & 0.108 \\ 
   &  & 350 & 315.500 & 0.033 & 266.000 & 1.000 & 100.500 & 0.133 \\ 
   &  & 1000 & 308.000 & 0.065 & 260.000 & 1.000 & 128.000 & 0.954 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Robinson-Fould distance from the null tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Bayesian_WXRFnull}
\end{table}