### 2022-03-24 Writing/Tables/Full_Tab_Bayesian_WXTrbest.tex

% latex table generated in R 4.1.3 by xtable 1.8-4 package
% Thu Mar 24 14:25:30 2022
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & minimised.maximised.W & minimised.maximised.p & minimised.randomised.W & minimised.randomised.p & maximised.randomised.W & maximised.randomised.p \\ 
  \hline
bayesian & 25 & 100 & 214.000 & 1.000 & 179.000 & 1.000 & 157.500 & 1.000 \\ 
   &  & 350 & 241.500 & 1.000 & 198.500 & 1.000 & 158.000 & 1.000 \\ 
   &  & 1000 & 158.500 & 1.000 & 96.000 & 0.092 & 165.000 & 1.000 \\ 
   & 75 & 100 & 101.000 & 0.121 & 143.000 & 1.000 & 235.000 & 1.000 \\ 
   &  & 350 & 143.000 & 1.000 & 137.000 & 1.000 & 180.000 & 1.000 \\ 
   &  & 1000 & 151.000 & 1.000 & 140.000 & 1.000 & 175.000 & 1.000 \\ 
   & 150 & 100 & 135.000 & 1.000 & 226.000 & 1.000 & 293.000 & 0.202 \\ 
   &  & 350 & 219.000 & 1.000 & 228.000 & 1.000 & 213.000 & 1.000 \\ 
   &  & 1000 & 178.000 & 1.000 & 170.000 & 1.000 & 196.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Triplets distance from the best tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Bayesian_WXTrbest}
\end{table}