### 2022-03-24 Writing/Tables/Full_Tab_Parsimony_WXTrbest.tex

% latex table generated in R 4.1.3 by xtable 1.8-4 package
% Thu Mar 24 14:25:30 2022
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & minimised.maximised.W & minimised.maximised.p & minimised.randomised.W & minimised.randomised.p & maximised.randomised.W & maximised.randomised.p \\ 
  \hline
parsimony & 25 & 100 & 194.500 & 1.000 & 208.000 & 1.000 & 207.000 & 1.000 \\ 
   &  & 350 & 184.000 & 1.000 & 199.500 & 1.000 & 199.000 & 1.000 \\ 
   &  & 1000 & 165.000 & 1.000 & 154.500 & 1.000 & 196.000 & 1.000 \\ 
   & 75 & 100 & 219.000 & 1.000 & 183.000 & 1.000 & 182.000 & 1.000 \\ 
   &  & 350 & 232.000 & 1.000 & 215.000 & 1.000 & 190.000 & 1.000 \\ 
   &  & 1000 & 186.500 & 1.000 & 221.500 & 1.000 & 225.000 & 1.000 \\ 
   & 150 & 100 & 205.000 & 1.000 & 193.000 & 1.000 & 211.000 & 1.000 \\ 
   &  & 350 & 220.000 & 1.000 & 242.000 & 1.000 & 233.000 & 1.000 \\ 
   &  & 1000 & 167.000 & 1.000 & 135.000 & 1.000 & 172.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Triplets distance from the best tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Parsimony_WXTrbest}
\end{table}