### 2022-03-24 Writing/Tables/Tab_pooledscharacters_test.tex

% latex table generated in R 4.1.3 by xtable 1.8-4 package
% Thu Mar 24 14:24:05 2022
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100 vs. c350 & 0.965 & 48263.500 & 0.000 \\ 
   & c100 vs. c1000 & 0.914 & 37714.500 & 0.000 \\ 
   & c350 vs. c1000 & 0.976 & 53138.000 & 0.000 \\ 
  Tr & c100 vs. c350 & 0.929 & 40253.000 & 0.000 \\ 
   & c100 vs. c1000 & 0.826 & 27923.000 & 0.000 \\ 
   & c350 vs. c1000 & 0.957 & 48164.500 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of characters: 100 (c100), 350 (c350) and 1000 (c1000). Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscharacters_test}
\end{table}