### 2022-05-16 Writing/Palaeontology/Tab_pooledscenarios_test.tex

% latex table generated in R 4.1.3 by xtable 1.8-4 package
% Thu Mar 24 14:24:05 2022
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & maximised vs. minimised & 0.848 & 97587.500 & \textbf{0.000} \\ 
   & maximised vs. randomised & \textbf{0.954} & 80382.500 & \textbf{0.000} \\ 
   & minimised vs. randomised & 0.941 & 44161.000 & \textbf{0.000} \\ 
  Tr & maximised vs. minimised & \textbf{0.992} & 68919.500 & 0.839 \\ 
   & maximised vs. randomised & \textbf{0.982} & 64222.500 & 1.000 \\ 
   & minimised vs. randomised & \textbf{0.980} & 60447.000 & 0.713 \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for each maximised, minimised and randomised character correlation. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscenarios_test}
\end{table}