### 2022-05-16 Writing/Tables/Tab_pooledstaxa_test.tex

% latex table generated in R 4.1.3 by xtable 1.8-4 package
% Thu Mar 24 14:24:05 2022
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & t25 vs. t75 & 0.949 & 79945.000 & 0.000 \\ 
   & t25 vs. t150 & 0.954 & 75906.000 & 0.000 \\ 
   & t75 vs. t150 & 0.984 & 60328.000 & 0.654 \\ 
  Tr & t25 vs. t75 & 0.960 & 72814.000 & 0.024 \\ 
   & t25 vs. t150 & 0.961 & 72716.000 & 0.027 \\ 
   & t75 vs. t150 & 0.980 & 64637.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of taxa: 25 (t25), 75 (t75) and 150 (t150). Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledstaxa_test}
\end{table}